---
title: "DDP vs ICE FY25 annual statistics"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

::: {.callout-note}
This report was generated using AI under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.
:::

```{r}
#| label: load-packages
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(knitr)
library(sf)
library(leaflet)
library(leafpop)
```

```{r}
#| label: load-data
source("R/ddp.R")
ddp_raw            <- tar_read(ddp_raw)
facilities_keyed   <- tar_read(facilities_keyed)
detloc_lookup      <- tar_read(detloc_lookup)
detloc_lookup_full <- tar_read(detloc_lookup_full)
vera_facilities    <- tar_read(vera_facilities)
```

```{r}
#| label: prepare-comparison
# Vera type lookup (one row per detloc).
# Uses corrected types from the pipeline (see vera_type_overrides() in vera-institute.R).
vera_type_lookup <- vera_facilities |>
  select(detloc, type_grouped = type_grouped_corrected,
         type_detailed = type_detailed_corrected, latitude, longitude) |>
  distinct(detloc, .keep_all = TRUE)

# ICE FY25 from pipeline
ice_fy25 <- facilities_keyed[["FY25"]] |>
  select(canonical_id, canonical_name, detloc, facility_name, facility_city,
         facility_state, sum_classification_levels) |>
  rename(ice_adp = sum_classification_levels)

ice_fy25_ids <- unique(ice_fy25$canonical_id)

# DDP facility-level summaries: Oct 1 2024 – Sep 24 2025
# (matches the ICE FY25 spreadsheet currency date)
ddp_fy25 <- ddp_facility_summary(ddp_raw,
                                  from = "2024-10-01", to = "2025-09-24") |>
  rename(ddp_adp = mean_pop)

# Use detloc_lookup_full to map DDP facility codes to canonical IDs.
# The 1:1 detloc_lookup picks only one code per canonical facility, but some
# facilities have multiple DETLOCs across sources (e.g., DMCP vs DDP codes).
# Using the full lookup ensures all codes contribute ADP to the right facility.
ddp_detloc_map <- detloc_lookup_full |>
  distinct(detloc, canonical_id)

ddp_fy25_keyed <- ddp_fy25 |>
  left_join(ddp_detloc_map, by = c("detention_facility_code" = "detloc")) |>
  mutate(in_ice_fy25 = !is.na(canonical_id) & canonical_id %in% ice_fy25_ids)
```

Throughout this report, facility-level summaries distinguish between facilities with **ADP ≥ 2** and those with **ADP < 2**. Facilities averaging fewer than two detainees per day---typically hospitals, occasional-use hold rooms, and other transient placements---account for many DDP facility codes but negligible population. Separating them keeps attention on facilities with substantive detention activity.

## Aggregate comparison

The DDP (Deportation Data Project) provides daily facility-level detention counts. To compare with the ICE FY25 annual statistics, we average DDP daily counts from October 1, 2024 through September 24, 2025---the date stamp on the FY25 spreadsheet.

```{r}
#| label: aggregate-totals
tibble(
  Source      = c("ICE FY25 annual stats", "DDP daily average"),
  Facilities  = c(nrow(ice_fy25), nrow(ddp_fy25)),
  `Total ADP` = c(round(sum(ice_fy25$ice_adp, na.rm = TRUE)),
                   round(sum(ddp_fy25$ddp_adp)))
) |>
  kable()
```

The DDP covers `r nrow(ddp_fy25)` facility codes while the ICE spreadsheet lists `r nrow(ice_fy25)` facilities. The DDP's higher total reflects the many smaller facilities (hold rooms, federal prisons, medical facilities) not included in the ICE annual statistics.

### Daily population over time

```{r}
#| label: fig-daily-total
#| fig-cap: "Total ICE detained population per day (DDP data, all facilities)"
#| fig-height: 4
daily_totals <- ddp_raw |>
  group_by(date) |>
  summarise(total_pop = sum(n_detained, na.rm = TRUE), .groups = "drop") |>
  mutate(fiscal_year = if_else(
    as.integer(format(date, "%m")) >= 10,
    paste0("FY", as.integer(format(date, "%Y")) + 1 - 2000),
    paste0("FY", as.integer(format(date, "%Y")) - 2000)
  ))

fy_breaks <- as.Date(c("2023-10-01", "2024-10-01", "2025-10-01"))
fy_labels <- data.frame(
  x = as.Date(c("2024-04-01", "2025-04-01", "2025-10-08")),
  label = c("FY24", "FY25", "FY26")
)

ggplot(daily_totals, aes(x = date, y = total_pop)) +
  geom_line(linewidth = 0.3) +
  geom_vline(xintercept = fy_breaks, linetype = "dashed", alpha = 0.4) +
  geom_text(data = fy_labels, aes(x = x, y = max(daily_totals$total_pop) * 1.04,
                                   label = label), inherit.aes = FALSE, size = 3) +
  scale_y_continuous(labels = comma) +
  labs(x = NULL, y = "Total detained population")
```

## Facility-level matching

We join the two datasets through canonical facility IDs. Each DDP facility code is mapped to a canonical ID via `detloc_lookup_full`, which preserves all DETLOC variants from every source (DMCP, DDP, hold/ERO, Vera). Some canonical facilities have multiple codes---for example, Central Arizona Florence Correctional Complex uses `CCADCAZ` (DMCP) and `CCAFLAZ` (DDP), and Delaney Hall uses `ESSEXNJ` (DMCP) and `DHDFNJ` (DDP). Using the full lookup ensures all codes contribute ADP to the correct facility.

```{r}
#| label: match-summary
# Aggregate DDP by canonical_id for matched facilities.
# Some canonical facilities have multiple DDP codes (e.g., Adelanto + Desert View Annex);
# summing by canonical_id captures the full DDP population for comparison.
ddp_by_canonical <- ddp_fy25_keyed |>
  filter(in_ice_fy25) |>
  group_by(canonical_id) |>
  summarise(ddp_adp = sum(ddp_adp), .groups = "drop")

comparison <- ice_fy25 |>
  # ICE FY25 can have multiple rows per canonical_id (e.g., Adelanto + Desert View Annex)
  group_by(canonical_id) |>
  summarise(
    canonical_name = first(canonical_name),
    ice_adp = sum(ice_adp, na.rm = TRUE),
    .groups = "drop"
  ) |>
  full_join(ddp_by_canonical, by = "canonical_id") |>
  mutate(
    in_ice = !is.na(ice_adp),
    in_ddp = !is.na(ddp_adp),
    diff   = ddp_adp - ice_adp,
    pct_diff = diff / ice_adp * 100
  )

matched <- comparison |> filter(in_ice & in_ddp)

tibble(
  Category = c("In both sources", "ICE only (no DDP match)",
               "Matched: ICE total ADP", "Matched: DDP total ADP",
               "Matched: aggregate difference"),
  Value = c(
    sum(comparison$in_ice & comparison$in_ddp),
    sum(comparison$in_ice & !comparison$in_ddp),
    round(sum(matched$ice_adp, na.rm = TRUE)),
    round(sum(matched$ddp_adp, na.rm = TRUE)),
    paste0(round(sum(matched$ddp_adp, na.rm = TRUE) -
                   sum(matched$ice_adp, na.rm = TRUE)),
           " (", round((sum(matched$ddp_adp, na.rm = TRUE) /
                           sum(matched$ice_adp, na.rm = TRUE) - 1) * 100, 2), "%)")
  )
) |>
  kable()
```

### Facility-level agreement

```{r}
#| label: fig-scatter
#| fig-cap: "FY25 ADP comparison by facility (log scale). Dashed line = perfect agreement."
#| fig-height: 5
both_nonzero <- matched |> filter(ice_adp > 0, ddp_adp > 0)

ggplot(both_nonzero, aes(x = ice_adp, y = ddp_adp)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_point(alpha = 0.6) +
  scale_x_log10(labels = comma) +
  scale_y_log10(labels = comma) +
  labs(x = "ICE FY25 ADP (annual stats)", y = "DDP mean daily population")
```

```{r}
#| label: agreement-bands
tibble(
  `Agreement band` = c("Within 10%", "Within 25%", "Within 50%", ">50% difference"),
  Facilities = c(
    sum(abs(both_nonzero$pct_diff) <= 10),
    sum(abs(both_nonzero$pct_diff) > 10 & abs(both_nonzero$pct_diff) <= 25),
    sum(abs(both_nonzero$pct_diff) > 25 & abs(both_nonzero$pct_diff) <= 50),
    sum(abs(both_nonzero$pct_diff) > 50)
  )
) |>
  kable()
```

Among the `r nrow(both_nonzero)` matched facilities with nonzero ADP in both sources, `r sum(abs(both_nonzero$pct_diff) <= 10)` agree within 10%. The log-scale correlation is `r round(cor(log(both_nonzero$ice_adp), log(both_nonzero$ddp_adp)), 3)`.

```{r}
#| label: tbl-ice-only
#| tbl-cap: "ICE FY25 facilities with no DDP match"
ice_only <- comparison |> filter(in_ice & !in_ddp)
if (nrow(ice_only) > 0) {
  ice_only |>
    select(canonical_id, canonical_name, ice_adp) |>
    mutate(ice_adp = round(ice_adp)) |>
    arrange(desc(ice_adp)) |>
    kable(col.names = c("ID", "Facility", "ICE ADP"))
}
```

`r if (nrow(ice_only) > 0) paste0(nrow(ice_only), " ICE FY25 facilities have no matching DDP facility code. These are cases where the DETLOC used in DMCP records does not appear in DDP data.") else "All ICE FY25 facilities have a matching DDP facility code."`

## Unmatched facilities {#sec-unmatched}

**`r sum(!ddp_fy25_keyed$in_ice_fy25)`** of `r nrow(ddp_fy25)` DDP facility codes do not appear in the ICE FY25 annual statistics. These are facilities where ICE holds detainees but which are excluded from the published spreadsheet.

```{r}
#| label: unmatched-setup
unmatched <- ddp_fy25_keyed |>
  filter(!in_ice_fy25) |>
  left_join(
    vera_type_lookup |> select(detloc, type_grouped, type_detailed),
    by = c("detention_facility_code" = "detloc")
  ) |>
  mutate(
    type_grouped = if_else(is.na(type_grouped), "Unclassified", type_grouped),
    adp_class = if_else(ddp_adp >= 2, "ADP \u2265 2", "ADP < 2")
  )
```

### By facility type

Facility types are drawn from the Vera Institute's classification system. "Dedicated" facilities are purpose-built or contractually dedicated to ICE detention. "Non-Dedicated" are county jails and other facilities that also hold ICE detainees. "Hold/Staging" are short-term ERO processing rooms. "Federal" includes BOP prisons and USMS facilities.

```{r}
#| label: tbl-unmatched-types
#| tbl-cap: "DDP facilities not in ICE FY25, by Vera facility type and ADP class"
unmatched |>
  group_by(type_grouped, adp_class) |>
  summarise(
    Facilities = n(),
    `FY25 ADP` = round(sum(ddp_adp)),
    .groups = "drop"
  ) |>
  arrange(desc(`FY25 ADP`)) |>
  kable(col.names = c("Facility type", "ADP class", "Facilities", "FY25 ADP"))
```

The `r nrow(unmatched)` unmatched facilities account for a full-year ADP of roughly **`r format(round(sum(unmatched$ddp_adp)), big.mark = ",")`**. Of those, `r sum(unmatched$adp_class == "ADP < 2")` facility codes average fewer than 2 detainees per day, contributing `r round(sum(unmatched$ddp_adp[unmatched$adp_class == "ADP < 2"]))` ADP in total.

```{r}
#| label: tbl-dedicated-unmatched
#| tbl-cap: "Dedicated facilities in DDP but absent from ICE FY25"
dedicated_unmatched <- unmatched |>
  filter(type_grouped == "Dedicated")

if (nrow(dedicated_unmatched) > 0) {
  dedicated_unmatched |>
    select(detention_facility_code, detention_facility, state, ddp_adp, adp_class) |>
    mutate(ddp_adp = round(ddp_adp)) |>
    arrange(desc(ddp_adp)) |>
    kable(col.names = c("DETLOC", "Facility name", "State", "FY25 ADP", "ADP class"))
}
```

`r if (nrow(dedicated_unmatched) > 0) paste0(nrow(dedicated_unmatched), " Dedicated facilities appear in DDP but not in ICE FY25, contributing ", round(sum(dedicated_unmatched$ddp_adp)), " ADP.") else "No Dedicated facilities are unmatched---all appear in the ICE FY25 spreadsheet."`

### Daily and monthly detention in unmatched facilities

```{r}
#| label: fig-daily-unmatched
#| fig-cap: "Daily detained population in facilities not in ICE FY25 annual statistics"
#| fig-height: 4
daily_unmatched <- ddp_raw |>
  filter(detention_facility_code %in% unmatched$detention_facility_code,
         date >= as.Date("2024-10-01"), date <= as.Date("2025-09-24")) |>
  group_by(date) |>
  summarise(total_pop = sum(n_detained, na.rm = TRUE), .groups = "drop")

ggplot(daily_unmatched, aes(x = date, y = total_pop)) +
  geom_line(linewidth = 0.3) +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  labs(x = NULL, y = "Detained population (unmatched facilities)")
```

```{r}
#| label: monthly-setup
unmatched_codes <- unmatched$detention_facility_code

monthly_unmatched <- ddp_raw |>
  filter(detention_facility_code %in% unmatched_codes,
         date >= as.Date("2024-10-01"), date <= as.Date("2025-09-24")) |>
  left_join(
    vera_type_lookup |> select(detloc, type_grouped),
    by = c("detention_facility_code" = "detloc")
  ) |>
  mutate(
    type_grouped = if_else(is.na(type_grouped), "Unclassified", type_grouped),
    month = as.Date(format(date, "%Y-%m-01"))
  )
```

```{r}
#| label: fig-monthly-total
#| fig-cap: "Average daily population in unmatched facilities, by month"
#| fig-height: 4
monthly_totals <- monthly_unmatched |>
  group_by(month) |>
  summarise(daily_avg = sum(n_detained) / n_distinct(date), .groups = "drop")

ggplot(monthly_totals, aes(x = month, y = daily_avg)) +
  geom_col() +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  labs(x = NULL, y = "Average daily population")
```

```{r}
#| label: fig-monthly-by-type
#| fig-cap: "Average daily population in unmatched facilities, by month and type"
#| fig-height: 5
monthly_by_type <- monthly_unmatched |>
  group_by(month, type_grouped) |>
  summarise(daily_avg = sum(n_detained) / n_distinct(date), .groups = "drop")

ggplot(monthly_by_type, aes(x = month, y = daily_avg, fill = type_grouped)) +
  geom_col() +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  labs(x = NULL, y = "Average daily population", fill = "Facility type")
```

### Last 60 days of FY25

The final 60 days of the FY25 data window (July 27 -- September 24, 2025) capture the period of peak detention.

```{r}
#| label: tbl-last60
#| tbl-cap: "Average daily population in unmatched facilities, last 60 days of FY25"
last60_start <- as.Date("2025-07-27")
last60_end   <- as.Date("2025-09-24")

last60 <- monthly_unmatched |>
  filter(date >= last60_start, date <= last60_end)

last60_by_type <- last60 |>
  group_by(type_grouped) |>
  summarise(
    Facilities = n_distinct(detention_facility_code),
    `Daily avg` = round(sum(n_detained) / n_distinct(date)),
    .groups = "drop"
  ) |>
  arrange(desc(`Daily avg`))

last60_total <- sum(last60_by_type$`Daily avg`)

last60_by_type |>
  rename(`Facility type` = type_grouped) |>
  kable()
```

In the last 60 days, **`r format(last60_total, big.mark = ",")`** people per day were held in facilities absent from the ICE FY25 annual statistics.

### Peak FY25 population by facility

```{r}
#| label: peak-setup
peak_fy25 <- ddp_facility_summary(ddp_raw,
                                   from = "2024-10-01", to = "2025-09-24",
                                   codes = unmatched_codes) |>
  mutate(
    mean_pop = round(mean_pop, 1),
    adp_class = if_else(mean_pop >= 2, "ADP \u2265 2", "ADP < 2")
  ) |>
  left_join(
    vera_type_lookup,
    by = c("detention_facility_code" = "detloc")
  ) |>
  mutate(type_grouped = if_else(is.na(type_grouped), "Unclassified", type_grouped))
```

```{r}
#| label: tbl-peak-top20
#| tbl-cap: "Top 20 unmatched facilities by FY25 peak detained population"
peak_fy25 |>
  arrange(desc(peak_pop)) |>
  head(20) |>
  select(detention_facility, state, type_grouped, peak_pop, peak_date, mean_pop) |>
  kable(col.names = c("Facility", "State", "Type", "Peak", "Peak date", "Mean ADP"))
```

```{r}
#| label: fig-peak-dist
#| fig-cap: "Distribution of FY25 peak population among unmatched facilities (log scale, facilities with peak > 0)"
#| fig-height: 4
peak_fy25 |>
  filter(peak_pop > 0) |>
  ggplot(aes(x = peak_pop)) +
  geom_histogram(bins = 40) +
  scale_x_log10(labels = comma) +
  labs(x = "Peak detained population (FY25)", y = "Facilities")
```

## Map of unmatched facilities

```{r}
#| label: fig-map
#| fig-cap: "DDP facilities not in ICE FY25 annual statistics, sized by FY25 peak population"
#| fig-height: 7
map_data <- peak_fy25 |>
  filter(peak_pop > 0, !is.na(latitude), !is.na(longitude))

map_sf <- st_as_sf(map_data, coords = c("longitude", "latitude"), crs = 4326)

type_colors <- c(
  "Dedicated"     = "#e41a1c",
  "Non-Dedicated" = "#377eb8",
  "Federal"       = "#4daf4a",
  "Hold/Staging"  = "#ff7f00",
  "Family/Youth"  = "#984ea3",
  "Medical"       = "#a65628",
  "Other/Unknown" = "#999999",
  "Unclassified"  = "#333333"
)

map_sf$type_grouped <- factor(map_sf$type_grouped, levels = names(type_colors))
present_levels <- levels(droplevels(map_sf$type_grouped))
pal <- colorFactor(palette = type_colors[present_levels], domain = present_levels)

# sqrt scale for area-proportional circle sizing
radii <- scales::rescale(sqrt(map_sf$peak_pop), to = c(3, 20))

leaflet(map_sf) |>
  addProviderTiles("CartoDB.Positron") |>
  addCircleMarkers(
    radius      = radii,
    color       = "black",
    weight      = 1,
    fillColor   = ~pal(type_grouped),
    fillOpacity = 0.7,
    label       = ~paste0(detention_facility, " (", state, ": peak ", peak_pop, ")"),
    popup       = ~paste0(
      "<b>", detention_facility, "</b><br>",
      state, " | ", type_grouped, "<br>",
      "Peak: ", peak_pop, " (", peak_date, ")<br>",
      "Mean ADP: ", mean_pop
    )
  ) |>
  addLegend(
    pal = pal, values = ~type_grouped,
    title = "Facility type", position = "bottomright"
  )
```

## Unreported detention: Biden final quarter vs early Trump {#sec-period-comparison}

The preceding sections establish which DDP facilities are absent from the ICE FY25 annual statistics. Here we compare the **unreported population**---detainees held at those unmatched facilities---during the final three months of the Biden administration versus a comparable window under the Trump administration.

```{r}
#| label: period-definitions
biden_from <- as.Date("2024-10-20")
biden_to   <- as.Date("2025-01-19")
trump_from <- as.Date("2025-07-17")
trump_to   <- as.Date("2025-10-15")
```

- **Biden final quarter**: `r biden_from` to `r biden_to` (92 days)
- **Trump recent quarter**: `r trump_from` to `r trump_to` (91 days)

```{r}
#| label: period-unmatched-summaries
# Summarise the unmatched facilities in each period
biden_unmatched <- ddp_facility_summary(ddp_raw,
                                         from = biden_from, to = biden_to,
                                         codes = unmatched_codes) |>
  left_join(vera_type_lookup |> select(detloc, type_grouped),
            by = c("detention_facility_code" = "detloc")) |>
  mutate(
    type_grouped = if_else(is.na(type_grouped), "Unclassified", type_grouped),
    adp_class = if_else(mean_pop >= 2, "ADP \u2265 2", "ADP < 2"),
    period = "Biden"
  )

trump_unmatched <- ddp_facility_summary(ddp_raw,
                                         from = trump_from, to = trump_to,
                                         codes = unmatched_codes) |>
  left_join(vera_type_lookup |> select(detloc, type_grouped),
            by = c("detention_facility_code" = "detloc")) |>
  mutate(
    type_grouped = if_else(is.na(type_grouped), "Unclassified", type_grouped),
    adp_class = if_else(mean_pop >= 2, "ADP \u2265 2", "ADP < 2"),
    period = "Trump"
  )
```

### Aggregate unreported population

```{r}
#| label: tbl-period-totals
#| tbl-cap: "Unreported DDP population by period (facilities not in ICE FY25)"
bind_rows(biden_unmatched, trump_unmatched) |>
  filter(peak_pop > 0) |>
  group_by(period, adp_class) |>
  summarise(
    Facilities = n(),
    `Total ADP` = round(sum(mean_pop)),
    .groups = "drop"
  ) |>
  arrange(adp_class, period) |>
  kable()
```

```{r}
#| label: tbl-period-headline
#| tbl-cap: "Unreported population totals (ADP \u2265 2 facilities only)"
bind_rows(biden_unmatched, trump_unmatched) |>
  filter(peak_pop > 0, adp_class == "ADP \u2265 2") |>
  group_by(period) |>
  summarise(
    Facilities = n(),
    `Total ADP` = round(sum(mean_pop)),
    `Mean ADP per facility` = round(sum(mean_pop) / n(), 1),
    .groups = "drop"
  ) |>
  kable()
```

### By facility type

```{r}
#| label: tbl-period-type-composition
#| tbl-cap: "Unreported facilities by type and period (ADP \u2265 2 only)"
type_comp <- bind_rows(biden_unmatched, trump_unmatched) |>
  filter(peak_pop > 0, adp_class == "ADP \u2265 2") |>
  group_by(period, type_grouped) |>
  summarise(
    Facilities = n(),
    ADP = round(sum(mean_pop)),
    .groups = "drop"
  )

type_comp |>
  pivot_wider(names_from = period,
              values_from = c(Facilities, ADP),
              values_fill = 0,
              names_glue = "{period} {.value}") |>
  mutate(
    `ADP change` = `Trump ADP` - `Biden ADP`,
    `% change` = if_else(
      `Biden ADP` == 0, NA_real_,
      round((`Trump ADP` / `Biden ADP` - 1) * 100, 1)
    )
  ) |>
  arrange(desc(`Trump ADP`)) |>
  kable(col.names = c("Facility type",
                       "Biden fac.", "Trump fac.",
                       "Biden ADP", "Trump ADP",
                       "ADP change", "% change"))
```

```{r}
#| label: tbl-period-type-transient
#| tbl-cap: "Transient facilities (ADP < 2) by type and period"
bind_rows(biden_unmatched, trump_unmatched) |>
  filter(peak_pop > 0, adp_class == "ADP < 2") |>
  group_by(period, type_grouped) |>
  summarise(
    Facilities = n(),
    ADP = round(sum(mean_pop), 1),
    .groups = "drop"
  ) |>
  pivot_wider(names_from = period,
              values_from = c(Facilities, ADP),
              values_fill = 0,
              names_glue = "{period} {.value}") |>
  arrange(desc(`Trump Facilities`)) |>
  kable(col.names = c("Facility type",
                       "Biden fac.", "Trump fac.",
                       "Biden ADP", "Trump ADP"))
```

### Facilities active in only one period

```{r}
#| label: period-exclusive-setup
biden_active <- biden_unmatched |> filter(peak_pop > 0, adp_class == "ADP \u2265 2")
trump_active <- trump_unmatched |> filter(peak_pop > 0, adp_class == "ADP \u2265 2")

biden_only_codes <- setdiff(biden_active$detention_facility_code,
                             trump_active$detention_facility_code)
trump_only_codes <- setdiff(trump_active$detention_facility_code,
                             biden_active$detention_facility_code)
```

```{r}
#| label: tbl-biden-only
#| tbl-cap: "Unreported facilities with ADP \u2265 2 active under Biden but not Trump"
if (length(biden_only_codes) > 0) {
  biden_unmatched |>
    filter(detention_facility_code %in% biden_only_codes) |>
    arrange(desc(mean_pop)) |>
    select(detention_facility, state, type_grouped, mean_pop, peak_pop) |>
    mutate(mean_pop = round(mean_pop, 1)) |>
    kable(col.names = c("Facility", "State", "Type", "Biden ADP", "Biden peak"))
}
```

`r length(biden_only_codes)` unreported facilities (ADP ≥ 2) were active in the Biden period but fell below the threshold or had zero population in the Trump period.

```{r}
#| label: tbl-trump-only
#| tbl-cap: "Unreported facilities with ADP \u2265 2 active under Trump but not Biden"
if (length(trump_only_codes) > 0) {
  trump_unmatched |>
    filter(detention_facility_code %in% trump_only_codes) |>
    arrange(desc(mean_pop)) |>
    select(detention_facility, state, type_grouped, mean_pop, peak_pop) |>
    mutate(mean_pop = round(mean_pop, 1)) |>
    kable(col.names = c("Facility", "State", "Type", "Trump ADP", "Trump peak"))
}
```

`r length(trump_only_codes)` unreported facilities (ADP ≥ 2) were active in the Trump period but not the Biden period.

```{r}
#| label: tbl-trump-only-by-type
#| tbl-cap: "New unreported facilities under Trump (ADP \u2265 2), by type"
if (length(trump_only_codes) > 0) {
  trump_unmatched |>
    filter(detention_facility_code %in% trump_only_codes) |>
    group_by(type_grouped) |>
    summarise(
      Facilities = n(),
      `Total ADP` = round(sum(mean_pop)),
      .groups = "drop"
    ) |>
    arrange(desc(`Total ADP`)) |>
    kable(col.names = c("Facility type", "Facilities", "Total ADP"))
}
```

### Facilities active in both periods

```{r}
#| label: both-periods-setup
both_codes <- intersect(biden_active$detention_facility_code,
                         trump_active$detention_facility_code)

both_compare <- biden_unmatched |>
  filter(detention_facility_code %in% both_codes) |>
  select(detention_facility_code, detention_facility, state, type_grouped,
         biden_adp = mean_pop, biden_peak = peak_pop) |>
  inner_join(
    trump_unmatched |>
      filter(detention_facility_code %in% both_codes) |>
      select(detention_facility_code,
             trump_adp = mean_pop, trump_peak = peak_pop),
    by = "detention_facility_code"
  ) |>
  mutate(
    adp_change = trump_adp - biden_adp,
    adp_pct_change = if_else(biden_adp == 0, NA_real_,
                             (trump_adp / biden_adp - 1) * 100)
  )
```

```{r}
#| label: fig-scatter-periods
#| fig-cap: "Mean ADP at unreported facilities: Biden vs Trump (facilities with ADP \u2265 2 in both periods, log scale)"
#| fig-height: 5
ggplot(both_compare, aes(x = biden_adp, y = trump_adp, color = type_grouped)) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_point(alpha = 0.7) +
  scale_x_log10(labels = comma) +
  scale_y_log10(labels = comma) +
  labs(x = "Biden final quarter ADP", y = "Trump recent quarter ADP",
       color = "Facility type")
```

```{r}
#| label: tbl-top-increases
#| tbl-cap: "Top 15 unreported facilities by ADP increase (Biden \u2192 Trump)"
both_compare |>
  arrange(desc(adp_change)) |>
  head(15) |>
  transmute(
    Facility = detention_facility, State = state, Type = type_grouped,
    `Biden ADP` = round(biden_adp),
    `Trump ADP` = round(trump_adp),
    Change = round(adp_change)
  ) |>
  kable()
```

```{r}
#| label: tbl-top-decreases
#| tbl-cap: "Top 15 unreported facilities by ADP decrease (Biden \u2192 Trump)"
both_compare |>
  arrange(adp_change) |>
  head(15) |>
  transmute(
    Facility = detention_facility, State = state, Type = type_grouped,
    `Biden ADP` = round(biden_adp),
    `Trump ADP` = round(trump_adp),
    Change = round(adp_change)
  ) |>
  kable()
```

### Daily unreported population by period

```{r}
#| label: fig-daily-unreported-periods
#| fig-cap: "Daily unreported detention population during both comparison periods, by facility type (ADP \u2265 2 facilities only)"
#| fig-height: 5
# Facility codes with ADP >= 2 in at least one period
substantive_codes <- union(biden_active$detention_facility_code,
                            trump_active$detention_facility_code)

daily_unreported <- ddp_raw |>
  filter(
    detention_facility_code %in% substantive_codes,
    (date >= biden_from & date <= biden_to) |
      (date >= trump_from & date <= trump_to)
  ) |>
  left_join(vera_type_lookup |> select(detloc, type_grouped),
            by = c("detention_facility_code" = "detloc")) |>
  mutate(
    type_grouped = if_else(is.na(type_grouped), "Unclassified", type_grouped),
    period = case_when(
      date >= biden_from & date <= biden_to ~ "Biden final quarter",
      date >= trump_from & date <= trump_to ~ "Trump recent quarter"
    )
  ) |>
  group_by(period, date, type_grouped) |>
  summarise(total_pop = sum(n_detained, na.rm = TRUE), .groups = "drop")

ggplot(daily_unreported, aes(x = date, y = total_pop, fill = type_grouped)) +
  geom_col() +
  facet_wrap(~period, scales = "free_x") +
  scale_y_continuous(labels = comma) +
  scale_x_date(date_labels = "%b %d") +
  labs(x = NULL, y = "Unreported detained population", fill = "Facility type")
```
