---
title: "DDP Daily Population Data — Parquet Summary & Integration"
date: last-modified
execute:
  echo: false
  warning: false
  message: false
---

## Overview

This document summarizes the Deportation Data Project (DDP) daily detention population dataset in parquet format (`data/ddp/facilities-daily-population-latest.parquet`). The parquet file is the current data source for FY26 and FY24 comparison analyses against ICE's published annual detention statistics.

The dataset spans October 2022 through March 2026. The first and last dates on which any facility recorded nonzero detention are **2022-10-01** and **2026-03-10**.

### Dataset Basics

| Metric | Value |
|--------|-------|
| **Format** | Apache Parquet |
| **Date range** | 2022-10-01 – 2026-03-10 |
| **Total observations** | 888,699 facility-day records |
| **Unique facility codes** | 707 |
| **Variables** | 9 |
| **Missing values** | None |

### Variables

| Column | Type | Description |
|--------|------|-------------|
| `detention_facility_code` | character | ICE DETLOC code |
| `date` | date | Observation date |
| `n_detained` | integer | Total detained population |
| `n_detained_at_midnight` | integer | Population at midnight |
| `n_detained_male` | integer | Male detainees |
| `n_detained_female` | integer | Female detainees |
| `n_detained_convicted_criminal` | integer | Detainees with criminal convictions |
| `n_detained_possibly_under_18` | integer | Detainees possibly under 18 |
| `detention_facility` | character | Facility name |

Unlike the earlier feather-format file, the parquet file does not include a `state` column. State is derived downstream via joins with `facility_roster` or `vera_facilities`.

---

## Population by Fiscal Year

```{r}
#| label: fy-summary
library(dplyr)

ddp_fy_summary <- tibble::tribble(
  ~`Fiscal Year`, ~`Days`,  ~`Mean Daily`, ~`Max Daily`, ~`Total Person-Days`,
  "FY2023",          365L,        41.1,       1776L,         10611760L,
  "FY2024",          366L,        54.6,       2203L,         14141343L,
  "FY2025",          365L,        70.7,       2940L,         18254136L,
  "FY2026 (partial)", 162L,       98.2,       4130L,         11175237L
)

knitr::kable(ddp_fy_summary, format.args = list(big.mark = ","))
```

Mean daily population is the aggregate across all facility codes. The median facility-day observation is zero, reflecting the large number of hold rooms, staging areas, and low-utilization codes in the data. FY2026 covers October 1, 2025 through March 10, 2026 only.

---

## Facility Code Coverage by Type

Type assignments use the project's unified type lookup: `detloc_lookup_full` joined to `facility_roster` (most authoritative), supplemented by Vera Institute `type_grouped_corrected` and `ddp_facility_canonical` for codes not in the main lookup.

```{r}
#| label: type-table
tibble::tribble(
  ~`Type`,           ~`Feather (853)`, ~`Parquet (707)`, ~`Both (620)`, ~`Feather-only (233)`, ~`Parquet-only (87)`,
  "Non-Dedicated",        209L,              242L,            206L,             3L,                  36L,
  "Hold/Staging",         181L,              187L,            181L,             0L,                   6L,
  "Federal",              171L,              183L,            170L,             1L,                  13L,
  "Dedicated",             51L,               54L,             51L,             0L,                   3L,
  "Family/Youth",           7L,                8L,              7L,             0L,                   1L,
  "Other/Unknown",          5L,                3L,              3L,             2L,                   0L,
  "Medical",              228L,                1L,              1L,           227L,                   0L,
  "Unclassified",           1L,               29L,              1L,             0L,                  28L
) |>
  knitr::kable()
```

The feather-only codes are almost entirely medical facilities (227 of 233). The parquet-only codes are predominantly non-dedicated jails (36), federal facilities (13), and hold/staging codes (6); 28 remain unclassified, meaning they have no match in any of the project's type reference files.

---

## Feather vs. Parquet Comparison

The feather file covers September 1, 2023 – October 15, 2025. The parquet file covers October 1, 2022 – March 10, 2026. Their overlap period is **September 1, 2023 – October 15, 2025** (776 days).

### Aggregate daily population

Over the overlap period, the parquet total is on average **209 detainees per day lower** than the feather total (feather mean: 44,607; parquet mean: 44,398). No two days are exactly equal. The largest single-day absolute difference is 1,773. The difference is consistent with the feather-only codes contributing approximately 60 ADP during this period — almost all from medical facilities. The parquet-only codes contributed approximately 1 ADP in total over the same period.

### Facility-level agreement

For the 620 codes present in both datasets, over the overlap period:

| Agreement band | Facilities |
|----------------|-----------|
| Exactly equal | 410 |
| Within 1% | 562 |
| Within 5% | 601 |
| Discrepant ≥ 5% | 19 |

Mean absolute difference for shared codes: 0.28 ADP.

The 19 discrepant facilities include several with very low ADP (< 2) where small absolute differences produce large percentages, and one substantive case: **Florida Soft-Sided Facility** (FLDSSFS), which shows 106 ADP in the feather vs. 26 ADP in the parquet over the overlap period.

---

## Data Integration Summary

### Pipeline Targets

The feather file remains the source for `ddp_raw` and all existing FY25-era pipeline targets. The parquet file is loaded as a separate `ddp_new` target for FY26 (and future FY24) comparisons:

```r
tar_target(ddp_parquet_file,
  here::here("data/ddp/facilities-daily-population-latest.parquet"),
  format = "file")

tar_target(ddp_new, arrow::read_parquet(ddp_parquet_file))
```

The `ddp_facility_summary()` function handles both sources; the `state` column it previously required is treated as optional.

### Canonical ID Matching

DDP facility codes are matched to canonical facility IDs via `detloc_lookup_full`. Match coverage for the 620 shared codes is the same as documented for the feather file. The 87 parquet-only codes are covered by the same lookup where they have known DETLOCs; 28 remain unclassified.

### Canonical ID Space

| ID Range | Count | Source |
|----------|-------|--------|
| 1–398 | ~347 | FY19–FY26 annual stats (frozen) |
| 1001–1053 | 53 | DMCP 2015/2017 only |
| 1054–1203 | 150 | DDP non-hold/non-medical |
| 2001–2025 | 25 | ERO field offices |
| 2026–2186 | 161 | Hold facilities |
| 3001–3226 | 226 | Medical facilities |

---

## FY26 Comparison Export

The `ddp_fy26_comparison_export` target exports 8 pre-computed RDS files to `data/ddp-comparison-export-fy26/` covering **October 1, 2025 – February 5, 2026** (the ICE data cutoff in the Feb 12 release):

| Object | Description |
|--------|-------------|
| `ice_fy26` | ICE FY26b spreadsheet: 220 facilities, total ADP 67,139 |
| `ddp_fy26` | DDP facility summary for the period (562 codes with peak > 0) |
| `ddp_fy26_keyed` | DDP summary with canonical IDs; `in_ice_fy26` flag |
| `daily_totals_fy26` | Aggregate daily population (128 days; peak 76,271 on 2026-01-16) |
| `unmatched_fy26` | 354 DDP-only facilities; 118 with ADP ≥ 2 |
| `daily_unmatched_fy26` | Daily population for unmatched codes |
| `monthly_unmatched_fy26` | Monthly breakdown by type for unmatched codes |
| `peak_fy26` | Peak summary with geocoding for unmatched facilities |

---

## Related Files

- `R/ddp.R` — `ddp_facility_summary()`, `build_fy26b()`, `export_ddp_fy26_comparison_data()`
- `data/ddp-comparison-export-fy26/` — pre-computed RDS files
- `facility_roster.csv` — canonical facility addresses, types, DETLOCs
- `ddp-integration-summary.md` — original feather integration notes
