---
title: "Detention List Wikipedia Tables"
format: html
---

```{r}
#| label: setup
#| include: false
library(targets)
library(dplyr)
source("R/wiki-match.R")
source("R/wikipedia.R")
source("R/gh-copy-assets.R")
fw <- tar_read(facilities_fy26_wiki)
fp <- tar_read(facility_presence)
panel <- tar_read(facilities_panel)
cwm <- tar_read(canonical_wiki_match)
```

Preview of 10 rows from each table variant, followed by a button to copy the full wikitext to clipboard.

## Standard wikitable

Links are validated against the live Wikipedia API via `check_wiki_articles()`. Only titles confirmed to exist become wikilinks; non-existent titles (parenthetical display names, red links in the source table) fall through to plain text.

```{r}
#| label: standard
#| results: asis
#| echo: false
wt_standard <- generate_fy26_wikitable(fw, facility_presence = fp)

# Preview first 10 data rows (skip header lines, take 10 row blocks)
wt_lines <- strsplit(wt_standard, "\n")[[1]]
# Header = lines up to first "|-\n| " data row; each data row starts with "|-"
row_markers <- which(grepl("^\\|-$", wt_lines))
# First row marker is before the header, data rows start from marker 2
n_preview <- min(10, length(row_markers) - 1)
# Keep everything up to the end of the 10th data row + closing tag
end_line <- if (n_preview < length(row_markers) - 1) {
  row_markers[n_preview + 2] - 1
} else {
  length(wt_lines)
}
preview_lines <- c(wt_lines[1:end_line], "|}")
cat("```\n")
cat(paste(preview_lines, collapse = "\n"))
cat("\n```\n")
```

```{r}
#| label: standard-copy
#| echo: false
include_table_assets()
copy_to_clipboard_button_wt(wt_standard, button_label = "Copy Wikitext (Standard)")
```


## All-links wikitable (with red links)

Every facility name is wikilinked. Facilities without a confirmed article get a bare `[[Name]]` link, which renders as a red link on Wikipedia.

```{r}
#| label: redlinks
#| results: asis
#| echo: false
wt_redlinks <- generate_fy26_wikitable(fw, facility_presence = fp,
                                       include_redlinks = TRUE)

wt_lines <- strsplit(wt_redlinks, "\n")[[1]]
row_markers <- which(grepl("^\\|-$", wt_lines))
n_preview <- min(10, length(row_markers) - 1)
end_line <- if (n_preview < length(row_markers) - 1) {
  row_markers[n_preview + 2] - 1
} else {
  length(wt_lines)
}
preview_lines <- c(wt_lines[1:end_line], "|}")
cat("```\n")
cat(paste(preview_lines, collapse = "\n"))
cat("\n```\n")
```

```{r}
#| label: redlinks-copy
#| echo: false
copy_to_clipboard_button_wt(wt_redlinks, button_label = "Copy Wikitext (All Links)")
```


## Red links only

Only facilities with **no** Wikipedia link (`wiki_slug` and `wiki_match` both missing). All names are wikilinked as red links.

```{r}
#| label: redlinks-only
#| results: asis
#| echo: false
fw_unlinked <- fw |>
  filter(is.na(wiki_slug) | wiki_slug == "",
         is.na(wiki_match) | wiki_match == "")

wt_redonly <- generate_fy26_wikitable(fw_unlinked, facility_presence = fp,
                                      include_redlinks = TRUE)

wt_lines <- strsplit(wt_redonly, "\n")[[1]]
row_markers <- which(grepl("^\\|-$", wt_lines))
n_preview <- min(10, length(row_markers) - 1)
end_line <- if (n_preview < length(row_markers) - 1) {
  row_markers[n_preview + 2] - 1
} else {
  length(wt_lines)
}
preview_lines <- c(wt_lines[1:end_line], "|}")
cat("```\n")
cat(paste(preview_lines, collapse = "\n"))
cat("\n```\n")
```

```{r}
#| label: redlinks-only-copy
#| echo: false
n_unlinked <- nrow(fw_unlinked)
n_total <- nrow(fw)
```

`r n_unlinked` of `r n_total` FY26 facilities have no Wikipedia link.

```{r}
#| label: redlinks-only-copy-btn
#| echo: false
copy_to_clipboard_button_wt(wt_redonly, button_label = "Copy Wikitext (Red Links Only)")
```


## Closed facilities

Facilities present in earlier fiscal years but absent from FY26. Status shows the first-seen to last-seen span. Wikipedia links use canonical-level matches where available.

```{r}
#| label: closed
#| results: asis
#| echo: false
wt_closed <- generate_closed_wikitable(panel, fp,
                                        canonical_wiki_match = cwm)

wt_lines <- strsplit(wt_closed, "\n")[[1]]
row_markers <- which(grepl("^\\|-$", wt_lines))
n_preview <- min(10, length(row_markers) - 1)
end_line <- if (n_preview < length(row_markers) - 1) {
  row_markers[n_preview + 2] - 1
} else {
  length(wt_lines)
}
preview_lines <- c(wt_lines[1:end_line], "|}")
cat("```\n")
cat(paste(preview_lines, collapse = "\n"))
cat("\n```\n")
```

```{r}
#| label: closed-copy
#| echo: false
copy_to_clipboard_button_wt(wt_closed, button_label = "Copy Wikitext (Closed Facilities)")
```


## Closed facilities (with red links)

Same as above, but every facility name is wikilinked.

```{r}
#| label: closed-redlinks
#| results: asis
#| echo: false
wt_closed_red <- generate_closed_wikitable(panel, fp,
                                            canonical_wiki_match = cwm,
                                            include_redlinks = TRUE)

wt_lines <- strsplit(wt_closed_red, "\n")[[1]]
row_markers <- which(grepl("^\\|-$", wt_lines))
n_preview <- min(10, length(row_markers) - 1)
end_line <- if (n_preview < length(row_markers) - 1) {
  row_markers[n_preview + 2] - 1
} else {
  length(wt_lines)
}
preview_lines <- c(wt_lines[1:end_line], "|}")
cat("```\n")
cat(paste(preview_lines, collapse = "\n"))
cat("\n```\n")
```

```{r}
#| label: closed-redlinks-copy
#| echo: false
copy_to_clipboard_button_wt(wt_closed_red, button_label = "Copy Wikitext (Closed, All Links)")
```
