---
title: "ERO Field Offices as Informal Detention Sites"
date: today
format: html
---

## Background

U.S. Immigration and Customs Enforcement (ICE) Enforcement and Removal Operations
(ERO) maintains 25 field offices nationwide. These offices have always contained
holding rooms — small, spartan spaces designed to hold people for a few hours
after arrest, pending transfer to an official detention facility.

In June 2025, ICE issued an internal policy memo waiving its prior 12-hour limit
on holds, extending permissible detention to up to three days. In practice, court
records and data analyzed by the Guardian show the agency has routinely exceeded
even this extended limit, holding people in unventilated concrete rooms for days
and sometimes weeks — without the inspection, legal access, or oversight
requirements that apply to official ICE detention centers.

This document harvests the 25 ERO field office listings from ICE.gov, builds a
structured dataset, and compares those locations against the holding facilities
named in the Guardian's October 2025 investigation.

---

## Scraping ERO Field Offices from ICE.gov

### Approach and limitations

The ICE field offices page (`https://www.ice.gov/contact/field-offices`) lists
offices of multiple types — ERO, HSI, OPLA, OPE, OPR — paginated 20 per page.
The page accepts query parameters to filter by office type (`?office=16` for ERO)
and by page (`?page=N`, zero-indexed).

In principle this is scrapable with `rvest`. In practice, a few complications
arise:

1. **Pagination**: the full ERO listing spans two pages.
2. **Inconsistent title/name split**: each listing has a short *title line*
   (e.g., `"Burlington - ERO"`) and a longer *official name* (e.g.,
   `"Boston Field Office"`). The city in the title is the physical address city;
   the official name sometimes refers to a different metropolitan area.
3. **Sparse fields**: phone numbers and addresses are inconsistently structured
   across entries; two offices (Dallas, St. Rose) have no phone listed.

The scraping approach below fetches both pages, extracts the structured office
blocks, and parses the title acronym as the `office_type`.

```{r}
#| label: scrape-setup
#| message: false

library(rvest)
library(dplyr)
library(stringr)
library(purrr)
library(tibble)
```

### Fetching the paginated ERO listings

The ERO filter URL returns 20 results on page 0 and 5 on page 1 (25 total,
matching ICE's own statement that ERO "operates 25 field offices nationwide").

```{r}
#| label: fetch-pages
#| eval: false   # set to true to re-fetch from ice.gov

library(xml2)
library(httr)
library(glue)

# 1. Correct the name and use the base URL
base_url <- "https://www.ice.gov/contact/field-offices"

fetch_ero_page <- function(page_num) {
  url <- glue("{base_url}?office=16&page={page_num}")
  
  # 2. Add a User-Agent header to look like a browser
  resp <- GET(
    url, 
    add_headers(`User-Agent` = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebkit/537.36")
  )
  
  # 3. Check for successful status before parsing
  if (status_code(resp) != 200) {
    warning(glue("Page {page_num} failed with status {status_code(resp)}"))
    return(NULL)
  }
  
  Sys.sleep(2) # Being a bit more polite for government servers
  read_html(content(resp, "text"))
}

# 4. Use map but safely (optional: filter out NULLs)
pages <- map(1:2, fetch_ero_page) %>% compact()
```

### Parsing office blocks

Each office on the page is rendered as a `<div>` with a title line and detail
block. The title line follows the pattern `"City - ACRONYM"`.

```{r}
#| label: parse-offices
#| eval: false

parse_ero_page <- function(html) {
  # Title lines are in <h2> or similar heading within each listing block;
  # detail text (address, phone, AOR) follows in paragraph elements.
  # NOTE: actual CSS selectors require inspection of the live DOM.
  # The extraction below is illustrative — adjust selectors as needed.

  blocks <- html |> html_elements(".views-row")

  map_dfr(blocks, function(b) {
    title       <- b |> html_element(".field-content h3, h2") |> html_text2()
    office_name <- b |> html_element(".views-field-title") |> html_text2()
    address_raw <- b |> html_element(".street-address, address") |> html_text2()
    phone       <- b |> html_elements("p") |> html_text2() |>
                        str_subset("^\\(\\d{3}\\)") |> first(default = NA_character_)
    aor_text    <- b |> html_elements("p, strong") |> html_text2() |>
                        str_subset("Area of Responsibility") |>
                        str_remove(".*Area of Responsibility:\\s*") |>
                        first(default = NA_character_)
    email_node  <- b |> html_element("a[href^='mailto']")
    email       <- if (!is.null(email_node)) html_attr(email_node, "href") |>
                        str_remove("^mailto:") else NA_character_

    tibble(
      listing_name  = trimws(title),
      office_type   = str_extract(title, "[A-Z]+$"),
      office_name   = trimws(office_name),
      address_raw   = trimws(address_raw),
      phone         = phone,
      aor           = aor_text,
      outreach_email = email
    )
  })
}

ero_scraped <- map_dfr(pages, parse_ero_page)
```

### Synthetic import

Because the ICE page does not use a simple table structure, and because CSS
selectors may shift between scrapes, the dataset below was assembled by hand
from the two fetched pages and verified against the live listings. This is the
authoritative version for downstream analysis.

```{r}
#| label: ero-field-offices

ero_field_offices <- tribble(
  ~listing_name,            ~office_type, ~office_name,                  ~address,                                 ~city,             ~state, ~zip,         ~phone,            ~area_of_responsibility,                                                                                           ~outreach_email,
  "Atlanta - ERO",          "ERO",        "Atlanta Field Office",         "180 Ted Turner Dr. SW, Suite 522",       "Atlanta",         "GA",   "30303",      "(404) 893-1290",  "Georgia, North Carolina, South Carolina",                                                                         "Atlanta.Outreach@ice.dhs.gov",
  "Baltimore - ERO",        "ERO",        "Baltimore Field Office",       "31 Hopkins Plaza, 6th Floor",            "Baltimore",       "MD",   "21201",      "(443) 560-0640",  "Maryland",                                                                                                        "Baltimore.Outreach@ice.dhs.gov",
  "Buffalo - ERO",          "ERO",        "Buffalo Field Office",         "250 Delaware Avenue, Floor 7",           "Buffalo",         "NY",   "14202",      "(716) 464-5800",  "Upstate New York",                                                                                                "Buffalo.Outreach@ice.dhs.gov",
  "Burlington - ERO",       "ERO",        "Boston Field Office",          "1000 District Avenue",                   "Burlington",      "MA",   "01803",      "(781) 359-7500",  "Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island, Vermont",                                          "Boston.Outreach@ice.dhs.gov",
  "Centennial - ERO",       "ERO",        "Denver Field Office",          "12445 E. Caley Avenue",                  "Centennial",      "CO",   "80111",      "(720) 873-2899",  "Colorado, Wyoming",                                                                                               "Denver.Outreach@ice.dhs.gov",
  "Chantilly - ERO",        "ERO",        "Washington Field Office",      "14797 Murdock Street",                   "Chantilly",       "VA",   "20598-5216", "(703) 633-2100",  "District of Columbia and Virginia",                                                                               "Washington.Outreach@ice.dhs.gov",
  "Chicago - ERO",          "ERO",        "Chicago Field Office",         "101 W Ida B Wells Drive, Suite 4000",    "Chicago",         "IL",   "60605",      "(872) 351-3990",  "Illinois, Indiana, Wisconsin, Missouri, Kentucky, Kansas",                                                         "Chicago.Outreach@ice.dhs.gov",
  "Dallas - ERO",           "ERO",        "Dallas Field Office",          "8101 N. Stemmons Frwy",                  "Dallas",          "TX",   "75247",      NA_character_,     "North Texas, Oklahoma",                                                                                            "Dallas.Outreach@ice.dhs.gov",
  "Detroit - ERO",          "ERO",        "Detroit Field Office",         "985 Michigan Avenue, Suite 207",         "Detroit",         "MI",   "48226",      "(313) 771-6601",  "Michigan, Ohio",                                                                                                  "Detroit.Outreach@ice.dhs.gov",
  "El Paso - ERO",          "ERO",        "El Paso Field Office",         "11541 Montana Ave, Suite E",             "El Paso",         "TX",   "79936",      "(915) 225-1901",  "West Texas, New Mexico",                                                                                          "ElPaso.Outreach@ice.dhs.gov",
  "Fort Snelling - ERO",    "ERO",        "St Paul Field Office",         "1 Federal Drive, Suite 1601",            "Fort Snelling",   "MN",   "55111",      "(612) 409-7799",  "Iowa, Minnesota, Nebraska, North Dakota, South Dakota",                                                            "StPaul.Outreach@ice.dhs.gov",
  "Harlingen - ERO",        "ERO",        "Harlingen Field Office",       "1717 Zoy Street",                        "Harlingen",       "TX",   "78552",      "(956) 389-7884",  "South Texas",                                                                                                     "Harlingen.Outreach@ice.dhs.gov",
  "Houston - ERO",          "ERO",        "Houston Field Office",         "126 Northpoint Drive",                   "Houston",         "TX",   "77060",      "(281) 774-4816",  "Southeast Texas",                                                                                                 "Houston.Outreach@ice.dhs.gov",
  "Los Angeles - ERO",      "ERO",        "Los Angeles Field Office",     "300 North Los Angeles St., Room 7631",   "Los Angeles",     "CA",   "90012",      "(213) 830-7911",  "LA Metro (LA, Orange, Riverside, San Bernardino); Central Coast (Ventura, Santa Barbara, San Luis Obispo)",     "LosAngeles.Outreach@ice.dhs.gov",
  "Miramar - ERO",          "ERO",        "Miramar Sub Office",           "2805 SW 145th Ave",                      "Miramar",         "FL",   "33027",      "(954) 843-5800",  "Florida, Puerto Rico, U.S. Virgin Islands",                                                                       "Miami.Outreach@ice.dhs.gov",
  "New York - ERO",         "ERO",        "New York City Field Office",   "26 Federal Plaza, 9th Floor",            "New York",        "NY",   "10278",      "(212) 436-9315",  "Five NYC boroughs; Duchess, Nassau, Putnam, Suffolk, Sullivan, Orange, Rockland, Ulster, Westchester counties",  "NewYork.Outreach@ice.dhs.gov",
  "Newark - ERO",           "ERO",        "Newark Field Office",          "970 Broad St., 11th Floor",              "Newark",          "NJ",   "07102",      "(862) 445-9200",  "New Jersey",                                                                                                      "Newark.Outreach@ice.dhs.gov",
  "Philadelphia - ERO",     "ERO",        "Philadelphia Field Office",    "114 North 8th Street",                   "Philadelphia",    "PA",   "19107",      "(215) 656-7164",  "Delaware, Pennsylvania, West Virginia",                                                                           "Philadelphia.Outreach@ice.dhs.gov",
  "Phoenix - ERO",          "ERO",        "Phoenix Field Office",         "2035 N. Central Avenue",                 "Phoenix",         "AZ",   "85004",      "(602) 257-5900",  "Arizona",                                                                                                         "Phoenix.Outreach@ice.dhs.gov",
  "San Antonio - ERO",      "ERO",        "San Antonio Field Office",     "1777 NE Loop 410, Floor 15",             "San Antonio",     "TX",   "78217",      "(210) 283-4750",  "Central Texas",                                                                                                   "SanAntonio.Outreach@ice.dhs.gov",
  "San Diego - ERO",        "ERO",        "San Diego Field Office",       "880 Front Street, #2242",                "San Diego",       "CA",   "92101",      "(619) 436-0410",  "San Diego and Imperial County",                                                                                   "SanDiego.Outreach@ice.dhs.gov",
  "San Francisco - ERO",    "ERO",        "San Francisco Field Office",   "630 Sansome Street, Rm 590",             "San Francisco",   "CA",   "94111",      "(415) 365-8800",  "Northern California, Hawaii, Guam, Saipan",                                                                       "SanFrancisco.Outreach@ice.dhs.gov",
  "Seattle - ERO",          "ERO",        "Seattle Field Office",         "12500 Tukwila International Boulevard",  "Seattle",         "WA",   "98168",      "(206) 277-2000",  "Alaska, Oregon, Washington",                                                                                      "Seattle.Outreach@ice.dhs.gov",
  "St. Rose - ERO",         "ERO",        "New Orleans Field Office",     "181 James Drive W",                      "St. Rose",        "LA",   "70087",      NA_character_,     "Alabama, Arkansas, Louisiana, Mississippi, Tennessee",                                                             "NewOrleans.Outreach@ice.dhs.gov",
  "West Valley City - ERO", "ERO",        "Salt Lake City Field Office",  "2975 Decker Lake Drive, Suite 100",      "West Valley City","UT",   "84119-6096", "(801) 736-1200",  "Utah, Idaho, Montana, Nevada",                                                                                    "SaltLakeCity.Outreach@ice.dhs.gov"
)
```

### Count

ICE's own page text states ERO operates 25 field offices. The listing confirms this:

```{r}
#| label: count-ero
nrow(ero_field_offices)
```

Two offices — Dallas and St. Rose (New Orleans) — have no phone number listed.
One entry (Miramar) is labeled a "Sub Office" rather than a "Field Office"; it is
the only ERO office of that designation. The Boston Field Office is listed under
Burlington, MA (its physical address), not Boston.

---

## Locations Named in the Guardian Investigation

The Guardian's October 2025 investigation
("Revealed: ICE violates its own policy by holding people in secretive rooms for
days or weeks", Olivares & Craft, 30 Oct 2025) named seven locations where
extended holding had been documented or litigated:

```{r}
#| label: guardian-cities

guardian_cities <- tribble(
  ~city,         ~state, ~guardian_notes,
  "New York",    "NY",   "10th floor, Jacob K Javits federal building; 600% avg detention increase post-June memo; one person held 2.5 months",
  "Baltimore",   "MD",   "6th floor, George H Fallon federal building; medical neglect; air mattresses added Mar 2025",
  "Los Angeles", "CA",   "Building basement; families with children held with little food or water (Guardian, Jun 2025)",
  "San Diego",   "CA",   "Courthouse basement (Times of San Diego, Oct 2025)",
  "Atlanta",     "GA",   "Agency office basement (Atlanta Journal-Constitution, Oct 2025)",
  "Santa Ana",   "CA",   "Congressional oversight denied (DC federal lawsuit)",
  "Washington",  "DC",   "Congressional oversight denied (DC federal lawsuit); DHS Secretary Noem denied ICE detains in field offices"
)
```

---

## Comparison: Guardian Locations vs. ERO Field Offices

```{r}
#| label: comparison

comparison <- guardian_cities |>
  left_join(
    ero_field_offices |> select(office_name, city, state, address),
    by = c("city", "state")
  ) |>
  mutate(
    ero_match = !is.na(office_name),
    discrepancy_note = case_when(
      city == "Santa Ana"  ~ "Orange County — within LA Field Office AOR but no separate ERO listing",
      city == "Washington" ~ "Washington Field Office address is in Chantilly, VA; DC holding facility is at a different location",
      TRUE ~ NA_character_
    )
  )

comparison |>
  select(city, state, ero_match, office_name, address, discrepancy_note)
```

Five of the seven Guardian locations match an ERO field office directly by
city and state. Two do not:

- **Santa Ana, CA** is within the Los Angeles Field Office's area of
  responsibility (which explicitly covers Orange County), but there is no
  dedicated ERO office listed there. This may be a sub-office or a non-ERO
  federal building.

- **Washington, DC** — the Washington Field Office is located in Chantilly, VA,
  and this is the facility referenced in the article. The city mismatch reflects
  the office's formal name rather than a separate DC location.

The Guardian's analysis further states that ICE has used "at least 170 ICE
holding facilities nationwide, including at 25 ICE field offices." That framing
suggests the 25 ERO field offices represent a subset of a much larger informal
detention infrastructure. The DDP data confirms this: it contains **173 distinct
hold codes**, including the 25 ERO field offices, 12 named sub-offices, and
approximately 130 local hold rooms spanning nearly every U.S. state.Seee ho

---

## Cross-check against canonical detention facilities

ERO field offices share cities with many official ICE detention facilities. Before
adding these as new facilities, we verify that no ERO office address duplicates
an existing canonical facility.

```{r}
#| label: address-crosscheck-setup
#| message: false

library(targets)
library(stringdist)

panel_facilities <- tar_read(panel_facilities)
```

```{r}
#| label: address-crosscheck

# Join ERO offices to canonical facilities by city/state
city_overlap <- ero_field_offices |>
  inner_join(
    panel_facilities |>
      select(canonical_id, canonical_name, facility_address, facility_city,
             facility_state, facility_zip),
    by = c("city" = "facility_city", "state" = "facility_state"),
    relationship = "many-to-many"
  ) |>
  mutate(
    addr_similarity = 1 - stringdist(
      tolower(address), tolower(facility_address),
      method = "osa"
    ) / pmax(nchar(address), nchar(facility_address))
  ) |>
  select(office_name, city, state, address, canonical_id, canonical_name,
         facility_address, addr_similarity) |>
  arrange(desc(addr_similarity))

addr_summary <- city_overlap |>
  group_by(office_name) |>
  slice_max(addr_similarity, n = 1, with_ties = FALSE) |>
  ungroup()

addr_summary |>
  select(office_name, address, canonical_name, facility_address, addr_similarity)
```

The highest address similarity across all city/state overlaps is
`r round(max(addr_summary$addr_similarity, na.rm = TRUE), 2)`. All matches are
between federal office buildings (ERO offices) and jails, processing centers,
hotels, or prisons — structurally different facilities. **No ERO field office
duplicates an existing canonical facility.**

```{r}
#| label: unmatched-cities

# ERO offices in cities with no existing canonical facility
ero_field_offices |>
  anti_join(
    panel_facilities |>
      select(facility_city, facility_state),
    by = c("city" = "facility_city", "state" = "facility_state")
  ) |>
  select(office_name, city, state)
```

---

## DDP DETLOC matching

The Deportation Data Project (DDP) daily population data contains 173 facility
codes ending in "HOLD" — these represent holding rooms at ICE offices,
sub-offices, and other locations. We match ERO field offices to their
corresponding DDP hold codes.

```{r}
#| label: ddp-setup
#| message: false

ddp_codes <- tar_read(ddp_codes)

hold_codes <- ddp_codes |>
  filter(grepl("HOLD", detention_facility_code, ignore.case = TRUE) |
         grepl("HOLD|HOLDING", detention_facility, ignore.case = TRUE))
```

### Matching methodology

For each ERO office, we search DDP hold codes in the same state where the city
name or the office name (e.g., "Boston" for the Burlington, MA office) appears
in the DDP facility name or code.

```{r}
#| label: ddp-matching

# Manual DETLOC assignments based on city/office-name matching and verification
ero_detloc_map <- tribble(
  ~office_name,                  ~detloc,    ~ddp_name,
  "Atlanta Field Office",        "ATLHOLD",  "Atlanta Dist. Hold Rm",
  "Baltimore Field Office",      "BALHOLD",  "Baltimore Hold Room",
  "Boston Field Office",         "BOSHOLD",  "Boston Hold Room",
  "Buffalo Field Office",        "BUFHOLD",  "Buffalo Hold Room",
  "Chicago Field Office",        "CHIHOLD",  "Chicago Hold Room",
  "Dallas Field Office",         "DALHOLD",  "Dallas F.o. Hold",
  "Denver Field Office",         "DENHOLD",  "Denver Hold Room",
  "Detroit Field Office",        "DETHOLD",  "Detroit Hold Room",
  "El Paso Field Office",        "IEPHOLD",  "El Paso Hold Room",
  "Harlingen Field Office",      "HLGHOLD",  "Harlingen Hold Room",
  "Houston Field Office",        "HOUHOLD",  "Houston Fo Holdroom",
  "Los Angeles Field Office",    "LOSHOLD",  "Los Cust Case",
  "Miramar Sub Office",          "MIAHOLD",  "Miami (Miramar) Hold Room",
  "New Orleans Field Office",    "NOLHOLD",  "New Orleans Hold Room",
  "New York City Field Office",  "NYCHOLD",  "Nyc Hold Room",
  "Newark Field Office",         "NEWHOLD",  "New/Ins Os Hold Room",
  "Philadelphia Field Office",   "PHIHOLD",  "Phi District Office",
  "Phoenix Field Office",        "PHOHOLD",  "Phoenix Dist Office",
  "Salt Lake City Field Office", "SLCHOLD",  "Salt Lake City Hold Room",
  "San Antonio Field Office",    "SNAHOLD",  "San Antonio Dro Hold Room",
  "San Diego Field Office",      NA_character_, NA_character_,
  "San Francisco Field Office",  "SFRHOLD",  "Sfr Hold Room",
  "Seattle Field Office",        "SEAHOLD",  "Seattle Field Office Hold Room",
  "St Paul Field Office",        "SPMHOLD",  "Bishop Henry Whipple Federal Bldg",
  "Washington Field Office",     "WASHOLD",  "Washington Field Office"
)

ero_detloc_map |>
  mutate(matched = !is.na(detloc)) |>
  count(matched)
```

```{r}
#| label: ddp-match-table

ero_detloc_map |>
  select(office_name, detloc, ddp_name)
```

**24 of 25** ERO offices match a DDP hold code. One office has no dedicated
hold code in the DDP data:

- **San Diego** — SNDHOLD is already canonical facility 327 ("San Diego
  District Staging") at a different address than the ERO field office at 880
  Front Street. No separate ERO hold code found.

---

## DDP hold room universe

Beyond the 25 ERO field offices, the DDP data contains a much larger network
of holding facilities. This section catalogs the full universe.

```{r}
#| label: hold-universe

hold_classified <- hold_codes |>
  mutate(
    facility_type = case_when(
      detention_facility_code %in% c(
        "ATLHOLD", "BALHOLD", "BOSHOLD", "BUFHOLD", "CHIHOLD", "DALHOLD",
        "DENHOLD", "DETHOLD", "HLGHOLD", "HOUHOLD", "LOSHOLD", "MIAHOLD",
        "NYCHOLD", "NEWHOLD", "NOLHOLD", "PHIHOLD", "PHOHOLD", "SNAHOLD",
        "IEPHOLD", "SFRHOLD", "SEAHOLD", "SPMHOLD", "SLCHOLD", "WASHOLD"
      ) ~ "ERO Field Office",
      grepl("sub.?office|sub-o", detention_facility, ignore.case = TRUE) ~
        "Sub-Office",
      grepl("field office|dist\\.? office|fo hold|district office",
            detention_facility, ignore.case = TRUE) ~
        "District/Satellite Office",
      grepl("staging|service staging", detention_facility, ignore.case = TRUE) ~
        "Staging Facility",
      grepl("ERO", detention_facility) ~ "ERO Hold Room",
      grepl("USBP|CBP|CPC|border", detention_facility, ignore.case = TRUE) ~
        "CBP Facility",
      grepl("command center", detention_facility, ignore.case = TRUE) ~
        "Command Center",
      grepl("custody case|cust case", detention_facility, ignore.case = TRUE) ~
        "Custody Case",
      TRUE ~ "Hold Room"
    )
  )

hold_classified |>
  count(facility_type) |>
  arrange(desc(n))
```

```{r}
#| label: sub-offices

hold_classified |>
  filter(facility_type == "Sub-Office") |>
  select(detention_facility_code, detention_facility, state) |>
  arrange(state)
```

The 12 named sub-offices are satellite operations within larger ERO field office
areas of responsibility (e.g., Anchorage under Seattle, Jacksonville under
Miami/Miramar, Honolulu under San Francisco).

```{r}
#| label: hold-by-state

hold_classified |>
  filter(facility_type == "Hold Room") |>
  count(state, name = "n_hold_rooms") |>
  arrange(desc(n_hold_rooms)) |>
  head(15)
```

The 122 generic hold rooms — local ICE offices with holding capacity — span
nearly every U.S. state and represent the broader network the Guardian
documented. Texas alone has `r hold_classified |> filter(facility_type == "Hold Room", state == "TX") |> nrow()` hold rooms.

---

## Canonical ID assignment

ERO field offices receive canonical IDs in the **2001–2025** range, reserved for
ICE, CBP, POE, and other DHS facilities used for detention.

```{r}
#| label: canonical-ids

ero_canonical <- ero_field_offices |>
  left_join(ero_detloc_map |> select(office_name, detloc), by = "office_name") |>
  arrange(office_name) |>
  mutate(
    canonical_id = 2000L + row_number(),
    canonical_name = office_name,
    facility_type_wiki = "ICE ERO Field Office"
  ) |>
  select(canonical_id, canonical_name, address, city, state, zip,
         detloc, facility_type_wiki, area_of_responsibility,
         phone, outreach_email)

ero_canonical |>
  select(canonical_id, canonical_name, city, state, detloc)
```

```{r}
#| label: save-canonical
readr::write_csv(
  ero_canonical,
  here::here("data", "ero-field-offices-canonical.csv")
)
```

Saved to `data/ero-field-offices-canonical.csv`.

---

## Wikipedia list table

A wikitable section for ERO field office holding facilities, formatted to match
the existing detention list article tables.

```{r}
#| label: wikitable-setup
#| message: false

source("R/wikipedia.R")
source("R/gh-copy-assets.R")
```

```{r}
#| label: wikitable-build

# Two city-link disambiguations (reuse existing overrides)
ero_wiki_df <- ero_canonical |>
  mutate(
    facility_name_wiki = canonical_name,
    city_state = paste0(city, ", ", state),
    location = paste0("[[", city_state, "]]"),
    status = "Active",
    facility_type_wiki = "ICE ERO Field Office",
    authority = "ERO",
    management = "ICE",
    average_daily_population = NA_character_,
    minimum_capacity = NA_character_,
    demographics = NA_character_
  ) |>
  select(facility_name_wiki, status, location, facility_type_wiki,
         authority, management, average_daily_population,
         minimum_capacity, demographics)

ero_wiki_df
```

```{r}
#| label: wikitable-generate
#| results: asis
#| echo: false

wt_ero <- generate_wikitable(
  ero_wiki_df,
  caption = "ICE ERO Field Office Holding Facilities",
  column_names = c(
    "Facility Name", "Status (year)", "Location", "Facility Type",
    "Authority", "Management", "Average Daily Population",
    "Minimum Capacity", "Demographics"
  )
)

wt_lines <- strsplit(wt_ero, "\n")[[1]]
row_markers <- which(grepl("^\\|-$", wt_lines))
n_preview <- min(10, length(row_markers) - 1)
end_line <- if (n_preview < length(row_markers) - 1) {
  row_markers[n_preview + 2] - 1
} else {
  length(wt_lines)
}
preview_lines <- c(wt_lines[1:end_line], "|}")
cat("```\n")
cat(paste(preview_lines, collapse = "\n"))
cat("\n```\n")
```

```{r}
#| label: wikitable-copy
#| echo: false
include_table_assets()
copy_to_clipboard_button_wt(wt_ero, button_label = "Copy Wikitext (ERO Field Offices)")
```
