---
title: "ICE Detention Facility Profiles: Before and After January 20, 2025"
subtitle: "DDP individual-level detention stints data (FOIA 2026-ICLI-00005)"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    # toc-float: true
    code-fold: true
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(dplyr)
library(tidyr)
library(knitr)

pd <- tar_read(facility_profiles_data)

# Cutoff display string
cutoff_str <- format(pd$cutoff, "%B %e, %Y")

# Friendly column-value labels used in multiple tables
criminality_order <- c(
  "1 Convicted Criminal",
  "2 Pending Criminal Charges",
  "3 Other Immigration Violator"
)
classification_order <- c("Low", "Medium / Low", "Medium / High", "High")
```

This report profiles five ICE detention facility groups, comparing detainee characteristics
**before** and **after** the Trump administration's inauguration on **`r cutoff_str`**.
Each profile is derived from individual-level detention stints data released by ICE via FOIA
and processed by the [Deportation Data Project](https://deportationdata.org).

A **stint** is one continuous period at one facility. People transferred between facilities
have a separate row for each location. Statistics here describe people *booked in* to
each facility before or after the cutoff date.

> **Note on period lengths:** The "Before" period covers October 2022 through January 19,
> 2025 (~2.3 years). The "After" period covers January 20, 2025 through March 11, 2026
> (~14 months). Raw stint counts therefore cannot be directly compared between periods;
> use the "Stints / year" and "Persons / year" figures in the Population tab for
> cross-period comparison. Percentage breakdowns within each period are directly comparable.

```{r facility-list}
pd$groups |>
  kable(
    col.names = c("Group key", "Full name / location", "DETLOC code(s)"),
    caption   = "Facility groups covered in this report"
  )
```

---

```{r helpers}
# Pivot one categorical summary (e.g. criminality) wide for a single facility
pivot_wide <- function(df, group_key, value_col = "value",
                       n_col = "n", pct_col = "pct",
                       levels = NULL) {
  df <- df |> filter(facility_group == group_key)
  if (!is.null(levels)) df <- df |> filter(.data[[value_col]] %in% levels)

  df |>
    select(period, value = all_of(value_col),
           n = all_of(n_col), pct = all_of(pct_col)) |>
    mutate(cell = paste0(n, " (", pct, "%)")) |>
    select(value, period, cell) |>
    pivot_wider(names_from = period, values_from = cell, values_fill = "—") |>
    rename(Category = value)
}

# Overview row for one facility, formatted for kable
fmt_overview <- function(g) {
  pd$overview |>
    filter(facility_group == g) |>
    transmute(
      Period          = as.character(period),
      Stints          = n_stints,
      `Median days`   = median_days,
      `Mean days`     = mean_days,
      `90th pct days` = p90_days,
      `% male`        = pct_male,
      `Median age`    = median_age,
      `% criminal+`   = pct_criminal,
      `% final order` = pct_final_order,
      `% removed`     = pct_removed,
      `% transferred` = pct_transferred
    )
}

# Nationality table for one facility: top 12, before | after, sorted by combined total
fmt_nationalities <- function(g) {
  grp <- pd$nationalities |> filter(facility_group == g)

  totals <- grp |>
    group_by(value) |>
    summarise(total = sum(n), .groups = "drop") |>
    arrange(desc(total))

  wide <- grp |>
    mutate(cell = paste0(n, " (", pct, "%)")) |>
    select(Nationality = value, period, cell) |>
    pivot_wider(names_from = period, values_from = cell, values_fill = "—")

  totals |>
    select(Nationality = value, total) |>
    left_join(wide, by = "Nationality") |>
    select(-total)
}

# Population / ADP table for one facility
fmt_population <- function(g) {
  pd$adp_and_totals |>
    filter(facility_group == g) |>
    transmute(
      Period              = as.character(period),
      `Period days`       = format(period_days, big.mark = ","),
      `ADP (stints)`      = adp_stints,
      `Stints booked in`  = format(n_stints,         big.mark = ","),
      `Unique persons`    = format(n_unique_persons,  big.mark = ","),
      `Stints / year`     = format(stints_per_year,   big.mark = ","),
      `Persons / year`    = format(persons_per_year,  big.mark = ",")
    )
}

# Render the DDP sparkline SVG (stored as character in pd$viz)
render_sparkline <- function(g) {
  svg <- pd$viz[[g]]$sparkline_svg
  if (is.null(svg) || nchar(svg) == 0) {
    cat("*No DDP daily population data for this facility.*\n")
  } else {
    cat(svg, "\n")
  }
}

# Render the ICE annual ADP bar chart SVG
render_adp_bars <- function(g) {
  svg <- pd$viz[[g]]$adp_bars_svg
  if (is.null(svg) || nchar(svg) == 0) {
    cat("*No ICE annual ADP data for this facility.*\n")
  } else {
    cat(svg, "\n")
  }
}
```

---

## IWAHOLD — AZ Removal Ops Coordination Center (AROCC)

*Phoenix AOR · Short-term staging and removal coordination facility*

::: {.panel-tabset}

### Population

```{r iwahold-pop}
fmt_population("IWAHOLD") |>
  kable(caption = "IWAHOLD: ADP and throughput before/after inauguration")
```

**ADP note:** "ADP (stints)" is computed from person-days (every overlapping stint clipped to the period window). "Stints booked in" and "Unique persons" count only arrivals during the period. "Per year" figures annualize by period length for cross-period comparison.

**ICE annual ADP (FY10–FY26):**

```{r iwahold-bars, results="asis"}
render_adp_bars("IWAHOLD")
```

**DDP daily population (Oct 2022 – Mar 2026):**

```{r iwahold-sparkline, results="asis"}
render_sparkline("IWAHOLD")
```

### Overview

```{r iwahold-overview}
fmt_overview("IWAHOLD") |>
  kable(caption = "IWAHOLD: summary statistics before/after inauguration")
```

### Nationalities

```{r iwahold-nat}
fmt_nationalities("IWAHOLD") |>
  kable(caption = "IWAHOLD: top nationalities (n, % of period stints)")
```

### Criminality

```{r iwahold-crim}
pivot_wide(pd$criminality, "IWAHOLD", levels = criminality_order) |>
  kable(caption = "IWAHOLD: book-in criminality classification")
```

### Security classification

```{r iwahold-class}
pivot_wide(pd$classification, "IWAHOLD", levels = classification_order) |>
  kable(caption = "IWAHOLD: detainee security classification")
```

### Threat level

```{r iwahold-threat}
pivot_wide(pd$threat_level, "IWAHOLD") |>
  kable(caption = "IWAHOLD: ICE case threat level")
```

### Release reason

```{r iwahold-release}
pivot_wide(pd$release_reason, "IWAHOLD") |>
  kable(caption = "IWAHOLD: stint release reason")
```

### Final order & mandatory detention

```{r iwahold-legal}
bind_rows(
  pivot_wide(pd$final_order, "IWAHOLD") |> mutate(Field = "Final order"),
  pivot_wide(pd$ina_236c,    "IWAHOLD") |> mutate(Field = "INA 236(c) mandatory")
) |>
  select(Field, Category, everything()) |>
  kable(caption = "IWAHOLD: final order status and mandatory detention")
```

### Age bands

```{r iwahold-age}
pivot_wide(pd$age_bands, "IWAHOLD", value_col = "age_band") |>
  kable(caption = "IWAHOLD: age distribution at book-in")
```

### Entry status

```{r iwahold-entry}
pivot_wide(pd$entry_status, "IWAHOLD") |>
  kable(caption = "IWAHOLD: immigration entry status")
```

:::

---

## SPMHOLD — Bishop Henry Whipple Federal Building (St. Paul)

*St. Paul AOR · Short-term ERO hold room; reflects Twin Cities arrest population*

::: {.panel-tabset}

### Population

```{r spmhold-pop}
fmt_population("SPMHOLD") |>
  kable(caption = "SPMHOLD: ADP and throughput before/after inauguration")
```

**ICE annual ADP (FY10–FY26):**

```{r spmhold-bars, results="asis"}
render_adp_bars("SPMHOLD")
```

**DDP daily population (Oct 2022 – Mar 2026):**

```{r spmhold-sparkline, results="asis"}
render_sparkline("SPMHOLD")
```

### Overview

```{r spmhold-overview}
fmt_overview("SPMHOLD") |>
  kable(caption = "SPMHOLD: summary statistics before/after inauguration")
```

### Nationalities

```{r spmhold-nat}
fmt_nationalities("SPMHOLD") |>
  kable(caption = "SPMHOLD: top nationalities (n, % of period stints)")
```

### Criminality

```{r spmhold-crim}
pivot_wide(pd$criminality, "SPMHOLD", levels = criminality_order) |>
  kable(caption = "SPMHOLD: book-in criminality classification")
```

### Security classification

```{r spmhold-class}
pivot_wide(pd$classification, "SPMHOLD", levels = classification_order) |>
  kable(caption = "SPMHOLD: detainee security classification")
```

### Threat level

```{r spmhold-threat}
pivot_wide(pd$threat_level, "SPMHOLD") |>
  kable(caption = "SPMHOLD: ICE case threat level")
```

### Release reason

```{r spmhold-release}
pivot_wide(pd$release_reason, "SPMHOLD") |>
  kable(caption = "SPMHOLD: stint release reason")
```

### Final order & mandatory detention

```{r spmhold-legal}
bind_rows(
  pivot_wide(pd$final_order, "SPMHOLD") |> mutate(Field = "Final order"),
  pivot_wide(pd$ina_236c,    "SPMHOLD") |> mutate(Field = "INA 236(c) mandatory")
) |>
  select(Field, Category, everything()) |>
  kable(caption = "SPMHOLD: final order status and mandatory detention")
```

### Age bands

```{r spmhold-age}
pivot_wide(pd$age_bands, "SPMHOLD", value_col = "age_band") |>
  kable(caption = "SPMHOLD: age distribution at book-in")
```

### Entry status

```{r spmhold-entry}
pivot_wide(pd$entry_status, "SPMHOLD") |>
  kable(caption = "SPMHOLD: immigration entry status")
```

:::

---

## PINEPLA — Pine Prairie ICE Processing Center

*New Orleans AOR · Dedicated long-term processing and removal center*

::: {.panel-tabset}

### Population

```{r pinepla-pop}
fmt_population("PINEPLA") |>
  kable(caption = "PINEPLA: ADP and throughput before/after inauguration")
```

**ICE annual ADP (FY10–FY26):**

```{r pinepla-bars, results="asis"}
render_adp_bars("PINEPLA")
```

**DDP daily population (Oct 2022 – Mar 2026):**

```{r pinepla-sparkline, results="asis"}
render_sparkline("PINEPLA")
```

### Overview

```{r pinepla-overview}
fmt_overview("PINEPLA") |>
  kable(caption = "PINEPLA: summary statistics before/after inauguration")
```

### Nationalities

```{r pinepla-nat}
fmt_nationalities("PINEPLA") |>
  kable(caption = "PINEPLA: top nationalities (n, % of period stints)")
```

### Criminality

```{r pinepla-crim}
pivot_wide(pd$criminality, "PINEPLA", levels = criminality_order) |>
  kable(caption = "PINEPLA: book-in criminality classification")
```

### Security classification

```{r pinepla-class}
pivot_wide(pd$classification, "PINEPLA", levels = classification_order) |>
  kable(caption = "PINEPLA: detainee security classification")
```

### Threat level

```{r pinepla-threat}
pivot_wide(pd$threat_level, "PINEPLA") |>
  kable(caption = "PINEPLA: ICE case threat level")
```

### Release reason

```{r pinepla-release}
pivot_wide(pd$release_reason, "PINEPLA") |>
  kable(caption = "PINEPLA: stint release reason")
```

### Final order & mandatory detention

```{r pinepla-legal}
bind_rows(
  pivot_wide(pd$final_order, "PINEPLA") |> mutate(Field = "Final order"),
  pivot_wide(pd$ina_236c,    "PINEPLA") |> mutate(Field = "INA 236(c) mandatory")
) |>
  select(Field, Category, everything()) |>
  kable(caption = "PINEPLA: final order status and mandatory detention")
```

### Age bands

```{r pinepla-age}
pivot_wide(pd$age_bands, "PINEPLA", value_col = "age_band") |>
  kable(caption = "PINEPLA: age distribution at book-in")
```

### Entry status

```{r pinepla-entry}
pivot_wide(pd$entry_status, "PINEPLA") |>
  kable(caption = "PINEPLA: immigration entry status")
```

:::

---

## DILLEY — South Texas Family Residential Center (STFRCTX + DILLSAF)

*San Antonio AOR · Family detention center; combined main unit and single-female wing*

::: {.panel-tabset}

### Population

```{r dilley-pop}
fmt_population("DILLEY") |>
  kable(caption = "DILLEY: ADP and throughput before/after inauguration")
```

**ICE annual ADP (FY10–FY26):**

```{r dilley-bars, results="asis"}
render_adp_bars("DILLEY")
```

**DDP daily population (Oct 2022 – Mar 2026):**

```{r dilley-sparkline, results="asis"}
render_sparkline("DILLEY")
```

### Overview

```{r dilley-overview}
fmt_overview("DILLEY") |>
  kable(caption = "DILLEY: summary statistics before/after inauguration")
```

### Nationalities

```{r dilley-nat}
fmt_nationalities("DILLEY") |>
  kable(caption = "DILLEY: top nationalities (n, % of period stints)")
```

### Criminality

```{r dilley-crim}
pivot_wide(pd$criminality, "DILLEY", levels = criminality_order) |>
  kable(caption = "DILLEY: book-in criminality classification")
```

### Security classification

```{r dilley-class}
pivot_wide(pd$classification, "DILLEY", levels = classification_order) |>
  kable(caption = "DILLEY: detainee security classification")
```

### Threat level

```{r dilley-threat}
pivot_wide(pd$threat_level, "DILLEY") |>
  kable(caption = "DILLEY: ICE case threat level")
```

### Release reason

```{r dilley-release}
pivot_wide(pd$release_reason, "DILLEY") |>
  kable(caption = "DILLEY: stint release reason")
```

### Final order & mandatory detention

```{r dilley-legal}
bind_rows(
  pivot_wide(pd$final_order, "DILLEY") |> mutate(Field = "Final order"),
  pivot_wide(pd$ina_236c,    "DILLEY") |> mutate(Field = "INA 236(c) mandatory")
) |>
  select(Field, Category, everything()) |>
  kable(caption = "DILLEY: final order status and mandatory detention")
```

### Age bands

```{r dilley-age}
pivot_wide(pd$age_bands, "DILLEY", value_col = "age_band") |>
  kable(caption = "DILLEY: age distribution at book-in")
```

### Entry status

```{r dilley-entry}
pivot_wide(pd$entry_status, "DILLEY") |>
  kable(caption = "DILLEY: immigration entry status")
```

:::

---

## NWDC — Northwest ICE Processing Center (Tacoma)

*Seattle AOR · Large dedicated processing center serving the Pacific Northwest*

::: {.panel-tabset}

### Population

```{r nwdc-pop}
fmt_population("NWDC") |>
  kable(caption = "NWDC: ADP and throughput before/after inauguration")
```

**ICE annual ADP (FY10–FY26):**

```{r nwdc-bars, results="asis"}
render_adp_bars("NWDC")
```

**DDP daily population (Oct 2022 – Mar 2026):**

```{r nwdc-sparkline, results="asis"}
render_sparkline("NWDC")
```

### Overview

```{r nwdc-overview}
fmt_overview("NWDC") |>
  kable(caption = "NWDC: summary statistics before/after inauguration")
```

### Nationalities

```{r nwdc-nat}
fmt_nationalities("NWDC") |>
  kable(caption = "NWDC: top nationalities (n, % of period stints)")
```

### Criminality

```{r nwdc-crim}
pivot_wide(pd$criminality, "NWDC", levels = criminality_order) |>
  kable(caption = "NWDC: book-in criminality classification")
```

### Security classification

```{r nwdc-class}
pivot_wide(pd$classification, "NWDC", levels = classification_order) |>
  kable(caption = "NWDC: detainee security classification")
```

### Threat level

```{r nwdc-threat}
pivot_wide(pd$threat_level, "NWDC") |>
  kable(caption = "NWDC: ICE case threat level")
```

### Release reason

```{r nwdc-release}
pivot_wide(pd$release_reason, "NWDC") |>
  kable(caption = "NWDC: stint release reason")
```

### Final order & mandatory detention

```{r nwdc-legal}
bind_rows(
  pivot_wide(pd$final_order, "NWDC") |> mutate(Field = "Final order"),
  pivot_wide(pd$ina_236c,    "NWDC") |> mutate(Field = "INA 236(c) mandatory")
) |>
  select(Field, Category, everything()) |>
  kable(caption = "NWDC: final order status and mandatory detention")
```

### Age bands

```{r nwdc-age}
pivot_wide(pd$age_bands, "NWDC", value_col = "age_band") |>
  kable(caption = "NWDC: age distribution at book-in")
```

### Entry status

```{r nwdc-entry}
pivot_wide(pd$entry_status, "NWDC") |>
  kable(caption = "NWDC: immigration entry status")
```

:::

---

## Data notes

- Source: [Deportation Data Project](https://deportationdata.org/data/processed/ice.html),
  FOIA release 2026-ICLI-00005 (data through March 11, 2026).
- **Before** = stints with `book_in_date_time` before January 20, 2025 (Oct 2022–Jan 19, 2025).
  **After** = stints booked in on or after January 20, 2025 (through March 11, 2026).
- Rows flagged as `likely_duplicate` (≈0.4% of all stints) are excluded.
- `case_threat_level` of "NA" in the source data means no ICE threat level assigned
  (typically non-criminal immigration violations), not a missing value.
- Nationality counts reflect `citizenship_country` as recorded by ICE. Top 12 per
  facility are shown (ranked by combined total across both periods).
- DILLEY combines DETLOC codes STFRCTX (South Texas Family Residential Center) and
  DILLSAF (Dilley Processing Single Female), both located in Dilley, TX.
