---
title: "ICE detention facility summary"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
execute:
  echo: true
  warning: false
  message: false
---

::: {.callout-note}
This report was generated using AI under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.

<!--
To indicate human review: Delete the line above about contents not being reviewed, and replace this comment with:
The contents have been reviewed and validated by [Your Name], [Your Role] on [Date].
-->
:::

```{r}
#| label: load-packages
library(targets)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(knitr)
```

```{r}
#| label: load-data
source_presence   <- tar_read(source_presence)
facility_presence <- tar_read(facility_presence)
facilities_panel  <- tar_read(facilities_panel)
hold_canonical    <- tar_read(hold_canonical_data)
ero_canonical     <- tar_read(ero_canonical)
```

## Overview

This report summarizes the ICE detention facility data assembled by the `targets` pipeline, covering `r length(unique(facilities_panel$fiscal_year))` fiscal years of data. The pipeline tracks facilities across multiple data sources and assigns each a canonical identifier.

```{r}
#| label: count-by-id-range
source_presence |>
  count(id_range) |>
  mutate(id_range = factor(id_range,
    levels = c("panel", "dmcp_only", "ddp_other", "ero", "hold", "medical"),
    labels = c("FY panel (1\u2013398)", "DMCP-only (1001\u20131053)",
               "DDP other (1054\u20131203)", "ERO offices (2001\u20132025)",
               "Hold facilities (2026\u20132186)", "Medical (3001\u20133226)")
  )) |>
  kable(col.names = c("ID range", "Facilities"))
```

The pipeline contains **`r nrow(source_presence)`** canonical facilities in total. Of these, **`r sum(source_presence$id_range == "panel")`** appear in the FY annual statistics panel (FY10--FY26, minus FY18).

### Active vs. closed (panel facilities)

```{r}
#| label: active-closed
most_recent <- tail(sort(unique(facilities_panel$fiscal_year)), 1)

facility_presence |>
  count(trajectory) |>
  arrange(desc(n)) |>
  kable(col.names = c("Trajectory", "Facilities"))
```

In `r most_recent`, **`r sum(facility_presence$last_seen == most_recent)`** panel facilities were active.

## Facility types

```{r}
#| label: types-active
active_types <- facilities_panel |>
  filter(fiscal_year == most_recent) |>
  count(facility_type_wiki, sort = TRUE) |>
  mutate(pct = n / sum(n))

active_types |>
  mutate(pct = percent(pct, accuracy = 0.1)) |>
  kable(col.names = c("Type", "Facilities", "Share"))
```

```{r}
#| label: fig-types-bar
#| fig-cap: !expr paste("Facility types among active", most_recent, "facilities")
#| fig-height: 4
ggplot(active_types, aes(x = n, y = reorder(facility_type_wiki, n))) +
  geom_col() +
  labs(x = "Facilities", y = NULL)
```

### Type distribution over time

```{r}
#| label: fig-types-time
#| fig-cap: "Facility type counts by fiscal year"
#| fig-height: 5
facilities_panel |>
  filter(!is.na(facility_type_wiki)) |>
  count(fiscal_year, facility_type_wiki) |>
  ggplot(aes(x = fiscal_year, y = n, fill = facility_type_wiki)) +
  geom_col() +
  labs(x = NULL, y = "Facilities", fill = "Type") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Source coverage

The pipeline draws on five data sources. The table below shows how many canonical facilities appear in each.

```{r}
#| label: source-coverage-summary
source_cols <- c("in_fy_panel", "in_faclist15", "in_faclist17",
                 "in_ddp", "in_marshall", "in_vera")
source_labels <- c("FY annual stats", "DMCP 2015 (faclist15)",
                   "DMCP 2017 (faclist17)", "DDP daily population",
                   "Marshall Project", "Vera Institute")

tibble(
  Source     = source_labels,
  Facilities = colSums(source_presence[source_cols]),
  Share      = percent(Facilities / nrow(source_presence), accuracy = 0.1)
) |>
  kable()
```

### Supplementary coverage

```{r}
#| label: supplementary-coverage
tibble(
  Attribute    = c("Has DETLOC code", "Is geocoded"),
  Facilities   = c(sum(source_presence$has_detloc),
                   sum(source_presence$is_geocoded)),
  Share        = percent(Facilities / nrow(source_presence), accuracy = 0.1)
) |>
  kable()
```

#### Geocode sources

The pipeline assembles geocoded coordinates from three sources: Google Maps API (panel and ERO facilities), The Marshall Project (hold rooms with known addresses), and the Vera Institute (backfill for hold rooms without Marshall coverage).

```{r}
#| label: geocode-source-breakdown
source_presence |>
  filter(is_geocoded) |>
  count(geocode_source, sort = TRUE) |>
  mutate(share = percent(n / sum(n), accuracy = 0.1)) |>
  kable(col.names = c("Geocode source", "Facilities", "Share"))
```

### Vera Institute coverage

```{r}
#| label: vera-not-canonical
vera_facilities <- tar_read(vera_facilities)
detloc_lookup   <- tar_read(detloc_lookup)

n_vera_total   <- n_distinct(vera_facilities$detloc)
n_vera_matched <- sum(unique(vera_facilities$detloc) %in% unique(detloc_lookup$detloc))
n_vera_outside <- n_vera_total - n_vera_matched

vera_outside <- vera_facilities |>
  filter(!detloc %in% unique(detloc_lookup$detloc)) |>
  count(type_grouped, sort = TRUE) |>
  mutate(share = percent(n / sum(n), accuracy = 0.1))

vera_outside |>
  kable(col.names = c("Vera facility type", "Facilities", "Share"))
```

The Vera Institute facility list contains **`r n_vera_total`** distinct facility codes, of which **`r n_vera_matched`** match a canonical facility DETLOC. The remaining **`r n_vera_outside`** Vera codes are not in the canonical dataset. These are predominantly jails and county facilities used for short-term detention ("Non-Dedicated"), medical facilities, and federal prisons not tracked in the ICE annual statistics.

### Sources per facility

```{r}
#| label: fig-n-sources
#| fig-cap: "Number of data sources per facility"
#| fig-height: 3.5
source_presence |>
  count(n_sources) |>
  ggplot(aes(x = factor(n_sources), y = n)) +
  geom_col() +
  geom_text(aes(label = n), vjust = -0.3, size = 3) +
  labs(x = "Number of data sources", y = "Facilities")
```

### Source overlap by ID range

```{r}
#| label: source-overlap-by-range
source_presence |>
  group_by(id_range) |>
  summarise(
    n          = n(),
    fy_panel   = sum(in_fy_panel),
    faclist15  = sum(in_faclist15),
    faclist17  = sum(in_faclist17),
    ddp        = sum(in_ddp),
    marshall   = sum(in_marshall),
    vera       = sum(in_vera),
    .groups = "drop"
  ) |>
  mutate(id_range = factor(id_range,
    levels = c("panel", "dmcp_only", "ddp_other", "ero", "hold", "medical"),
    labels = c("Panel", "DMCP-only", "DDP other", "ERO", "Hold", "Medical")
  )) |>
  kable(col.names = c("ID range", "Total", "FY panel", "faclist15",
                       "faclist17", "DDP", "Marshall", "Vera"))
```

## Source overlap analysis

The pipeline draws on four broad data source families. This section consolidates the two ICE-published sources (FY19–26 annual statistics and FY10–17 DMCP summary listings) into a single "ICE" category and compares overlap with DDP, Vera Institute, and Marshall Project.

```{r}
#| label: source-overlap-prep
library(ggVennDiagram)

sp_overlap <- source_presence |>
  mutate(in_ice = in_fy_panel | in_faclist15 | in_faclist17)

source_sets <- list(
  "ICE" = sp_overlap$canonical_id[sp_overlap$in_ice],
  "DDP" = sp_overlap$canonical_id[sp_overlap$in_ddp],
  "Vera" = sp_overlap$canonical_id[sp_overlap$in_vera],
  "Marshall" = sp_overlap$canonical_id[sp_overlap$in_marshall]
)
```

### Venn diagram

```{r}
#| label: fig-venn-diagram
#| fig-cap: "Source coverage across canonical facilities"
#| fig-height: 5.5
#| fig-width: 6
ggVennDiagram(source_sets, label_alpha = 0) +
  scale_fill_gradient(low = "white", high = "#4292C6") +
  theme(legend.position = "none")
```

### Pairwise overlap

```{r}
#| label: pairwise-overlap
src_names <- names(source_sets)
pair_mat <- matrix(NA_integer_, length(src_names), length(src_names),
                   dimnames = list(src_names, src_names))
for (i in seq_along(src_names)) {
  for (j in seq_along(src_names)) {
    pair_mat[i, j] <- length(intersect(source_sets[[i]], source_sets[[j]]))
  }
}
kable(pair_mat)
```

### Intersection patterns

```{r}
#| label: intersection-patterns
sp_patterns <- sp_overlap |>
  count(in_ice, in_ddp, in_vera, in_marshall, sort = TRUE) |>
  mutate(
    pattern = paste0(
      if_else(in_ice,      "ICE", ""),
      if_else(in_ddp,      " + DDP", ""),
      if_else(in_vera,     " + Vera", ""),
      if_else(in_marshall,  " + Marshall", "")
    ) |> stringr::str_remove("^\\s\\+\\s") |> trimws(),
    n_sources = in_ice + in_ddp + in_vera + in_marshall
  ) |>
  arrange(desc(n))

sp_patterns |>
  select(pattern, n_sources, n) |>
  kable(col.names = c("Sources present", "Count", "Facilities"))
```

```{r}
#| label: fig-intersection-upset
#| fig-cap: "Facility counts by source combination"
#| fig-height: 5
sp_patterns |>
  filter(n >= 1, pattern != "") |>
  mutate(pattern = forcats::fct_reorder(pattern, n)) |>
  ggplot(aes(x = n, y = pattern)) +
  geom_col() +
  geom_text(aes(label = n), hjust = -0.2, size = 3) +
  labs(x = "Number of facilities", y = NULL) +
  scale_x_continuous(expand = expansion(mult = c(0, 0.15)))
```

**`r sum(sp_overlap$in_ice & sp_overlap$in_ddp & sp_overlap$in_vera & sp_overlap$in_marshall)`** facilities (`r scales::percent(sum(sp_overlap$in_ice & sp_overlap$in_ddp & sp_overlap$in_vera & sp_overlap$in_marshall) / nrow(sp_overlap), accuracy = 0.1)`) appear in all four source families. **`r sum(!sp_overlap$in_ice & sp_overlap$in_ddp & sp_overlap$in_vera & sp_overlap$in_marshall)`** appear in DDP + Vera + Marshall but not any ICE-published source — these are primarily hold rooms and staging facilities that ICE does not include in its published statistics. **`r sum(sp_overlap$in_ice & !sp_overlap$in_ddp & !sp_overlap$in_vera & !sp_overlap$in_marshall)`** facilities appear only in ICE-published data with no external corroboration; these are mostly closed or short-lived sites with no DETLOC code.

## Coverage gaps

Panel facilities (IDs 1--398) with limited source coverage:

```{r}
#| label: coverage-gaps
panel_gaps <- source_presence |>
  filter(id_range == "panel") |>
  filter(n_sources <= 1) |>
  select(canonical_id, canonical_name, in_fy_panel, in_faclist15,
         in_faclist17, in_ddp, in_marshall, in_vera, has_detloc) |>
  arrange(canonical_id)

panel_gaps |>
  kable(col.names = c("ID", "Name", "FY panel", "FL15", "FL17",
                       "DDP", "Marshall", "Vera", "DETLOC"))
```

**`r nrow(panel_gaps)`** panel facilities appear in only one data source. Of these, **`r sum(!panel_gaps$has_detloc)`** have no DETLOC code, limiting the ability to link them across datasets.
