---
title: "Geocoding divergence report"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    page-layout: article
execute:
  echo: true
  warning: false
  message: false
---

::: {.callout-note}
This report was generated using AI under general human direction. At the time of generation, the contents have not been comprehensively reviewed by a human analyst.

<!--
To indicate human review: Delete the line above about contents not being reviewed, and replace this comment with:
The contents have been reviewed and validated by [Your Name], [Your Role] on [Date].
-->
:::

```{r}
#| label: load-packages
library(targets)
library(dplyr)
library(sf)
library(leaflet)
library(leaflet.extras)
library(htmltools)
library(knitr)
```

```{r}
#| label: load-data
# Source R/ functions so geocode_source_preference() is available
tar_source()
facilities_geocoded_all <- tar_read(facilities_geocoded_all)
```

## Summary

```{r}
#| label: summary-geocode-source
facilities_geocoded_all |>
  count(geocode_source, name = "facilities") |>
  arrange(desc(facilities)) |>
  kable(col.names = c("Geocode source", "Facilities"))
```

```{r}
#| label: summary-address-quality
facilities_geocoded_all |>
  count(address_quality, name = "facilities") |>
  arrange(desc(facilities)) |>
  kable(col.names = c("Address quality", "Facilities"))
```

```{r}
#| label: summary-divergent-by-source
facilities_geocoded_all |>
  filter(divergent) |>
  count(source_name, name = "divergent_facilities") |>
  arrange(desc(divergent_facilities)) |>
  kable(col.names = c("Source", "Divergent facilities"))
```

## Divergence map

Each line segment connects the Google Maps geocode (blue) to the source-provided coordinate (red) for facilities where the two locations differ by more than 1 km. The map is searchable by facility name, city, or state.

```{r}
#| label: build-map-data
divergent <- facilities_geocoded_all |>
  filter(divergent) |>
  mutate(
    divergence_km = round(divergence_km, 1),
    search_label = paste0(canonical_name, " (", facility_city, ", ", facility_state, ")"),
    popup_html = paste0(
      "<b>", htmlEscape(canonical_name), "</b> (ID ", canonical_id, ")<br>",
      ifelse(is.na(facility_address), "", paste0(htmlEscape(facility_address), "<br>")),
      htmlEscape(facility_city), ", ", facility_state,
      ifelse(is.na(facility_zip), "", paste0(" ", facility_zip)),
      "<br>Source: ", source_name,
      "<br>Divergence: ", divergence_km, " km"
    )
  )

# Read source preferences from the pipeline's single source of truth
prefs <- geocode_source_preference()

divergent <- divergent |>
  left_join(prefs, by = "canonical_id") |>
  mutate(
    recommendation = case_when(
      preferred_source == "google_maps"      ~ "rec: Google",
      preferred_source == "vera_institute"   ~ "rec: Vera/source",
      preferred_source == "marshall_project" ~ "rec: Marshall",
      preferred_source == "manual"           ~ "rec: manual",
      TRUE                                   ~ "unreviewed"
    )
  ) |>
  select(-preferred_source)
```

::: {.column-page-inset}

```{r}
#| label: divergence-map
#| fig-height: 8
# Build line segments
line_features <- lapply(seq_len(nrow(divergent)), function(i) {
  row <- divergent[i, ]
  st_linestring(matrix(
    c(row$google_lon, row$source_lon, row$google_lat, row$source_lat),
    ncol = 2
  ))
}) |> st_sfc(crs = 4326)

google_pts <- divergent |> filter(!is.na(google_lat))
source_pts <- divergent |> filter(!is.na(source_lat))

# Recommendation subsets
rec_google_d   <- divergent |> filter(recommendation == "rec: Google")
rec_vera_d     <- divergent |> filter(recommendation == "rec: Vera/source")
rec_marshall_d <- divergent |> filter(recommendation == "rec: Marshall")

# Helper: build line sfc from a divergent subset
make_lines <- function(df) {
  lapply(seq_len(nrow(df)), function(i) {
    row <- df[i, ]
    st_linestring(matrix(
      c(row$google_lon, row$source_lon, row$google_lat, row$source_lat),
      ncol = 2
    ))
  }) |> st_sfc(crs = 4326)
}

# Helper: add a recommendation layer group (line + preferred point + grey rejected point)
add_rec_layer <- function(map, df, group, preferred_color) {
  if (nrow(df) == 0) return(map)
  prefer_google <- group == "Rec: Google"

  map |>
    # Connecting line
    addPolylines(
      data = st_sf(df, geometry = make_lines(df)),
      color = preferred_color, weight = 2, opacity = 0.5,
      group = group, popup = ~popup_html
    ) |>
    # Preferred point (colored)
    addCircleMarkers(
      data = df,
      lng = if (prefer_google) ~google_lon else ~source_lon,
      lat = if (prefer_google) ~google_lat else ~source_lat,
      radius = 7, color = preferred_color, fillColor = preferred_color,
      fillOpacity = 0.9, weight = 2, opacity = 0.9,
      group = group, label = ~search_label, popup = ~popup_html
    ) |>
    # Rejected point (grey)
    addCircleMarkers(
      data = df,
      lng = if (prefer_google) ~source_lon else ~google_lon,
      lat = if (prefer_google) ~source_lat else ~google_lat,
      radius = 5, color = "#999", fillColor = "#bbb",
      fillOpacity = 0.6, weight = 1, opacity = 0.6,
      group = group, label = ~search_label, popup = ~popup_html
    )
}

m <- leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  # Line segments connecting Google ↔ source for each facility
  addPolylines(
    data = st_sf(divergent, geometry = line_features),
    color = "#333", weight = 3, opacity = 0.7,
    group = "Divergence lines",
    popup = ~popup_html
  ) |>
  # Google points (blue)
  addCircleMarkers(
    data = google_pts,
    lng = ~google_lon, lat = ~google_lat,
    radius = 5, color = "steelblue", fillColor = "steelblue",
    fillOpacity = 0.8, weight = 1, opacity = 0.8,
    group = "Google",
    label = ~search_label,
    popup = ~popup_html
  ) |>
  # Source points (red)
  addCircleMarkers(
    data = source_pts,
    lng = ~source_lon, lat = ~source_lat,
    radius = 5, color = "firebrick", fillColor = "firebrick",
    fillOpacity = 0.8, weight = 1, opacity = 0.8,
    group = "Source",
    label = ~search_label,
    popup = ~popup_html
  ) |>
  # Recommendation review layers (each shows line + preferred + grey rejected)
  add_rec_layer(rec_google_d,   "Rec: Google",   "steelblue") |>
  add_rec_layer(rec_vera_d,     "Rec: Vera",     "firebrick") |>
  add_rec_layer(rec_marshall_d, "Rec: Marshall",  "#e90") |>
  addLayersControl(
    overlayGroups = c("Divergence lines", "Google", "Source",
                      "Rec: Google", "Rec: Vera", "Rec: Marshall"),
    options = layersControlOptions(collapsed = FALSE)
  ) |>
  # Hide recommendation layers by default
  hideGroup("Rec: Google") |>
  hideGroup("Rec: Vera") |>
  hideGroup("Rec: Marshall") |>
  addSearchFeatures(
    targetGroups = c("Google", "Source"),
    options = searchFeaturesOptions(
      zoom = 9,
      openPopup = TRUE,
      position = "topleft",
      hideMarkerOnCollapse = TRUE
    )
  )

m
```

:::

## Operational significance of divergent facilities

How much does geocoding divergence matter? This section classifies divergent facilities by their presence in the FY26 annual stats panel and/or DDP daily population data, then summarizes peak ADP to gauge operational significance.

```{r}
#| label: divergent-presence
# Classify divergent facilities by data presence
facility_presence <- tar_read(facility_presence)
source_presence <- tar_read(source_presence)
facilities_panel <- tar_read(facilities_panel)
ddp_fy25_summary <- tar_read(ddp_fy25_summary)
detloc_lookup <- tar_read(detloc_lookup)

# FY26 presence from facility_presence (panel IDs 1-398)
fy26_ids <- facility_presence |>
  filter(FY26) |>
  pull(canonical_id)

# DDP presence from source_presence
ddp_ids <- source_presence |>
  filter(in_ddp) |>
  pull(canonical_id)

# Peak ADP from panel data (max across all fiscal years)
peak_adp_panel <- facilities_panel |>
  group_by(canonical_id) |>
  summarise(peak_adp_panel = max(adp, na.rm = TRUE), .groups = "drop") |>
  mutate(peak_adp_panel = ifelse(is.infinite(peak_adp_panel), NA_real_, peak_adp_panel))

# DDP ADP via detloc_lookup join
ddp_adp <- ddp_fy25_summary |>
  inner_join(detloc_lookup, by = c("detention_facility_code" = "detloc")) |>
  group_by(canonical_id) |>
  summarise(
    ddp_adp = max(adp_total, na.rm = TRUE),
    ddp_peak = max(peak_population, na.rm = TRUE),
    .groups = "drop"
  )

divergent_classified <- divergent |>
  mutate(
    in_fy26 = canonical_id %in% fy26_ids,
    in_ddp = canonical_id %in% ddp_ids,
    presence = case_when(
      in_fy26 & in_ddp  ~ "Both FY26 + DDP",
      in_fy26 & !in_ddp ~ "FY26 only",
      !in_fy26 & in_ddp ~ "DDP only",
      TRUE               ~ "Neither"
    )
  ) |>
  left_join(peak_adp_panel, by = "canonical_id") |>
  left_join(ddp_adp, by = "canonical_id") |>
  mutate(
    # Best available ADP: panel peak, then DDP peak
    best_peak_adp = coalesce(peak_adp_panel, ddp_peak)
  )
```

```{r}
#| label: presence-summary
divergent_classified |>
  group_by(presence) |>
  summarise(
    n = n(),
    median_peak_adp = round(median(best_peak_adp, na.rm = TRUE)),
    max_peak_adp = round(max(best_peak_adp, na.rm = TRUE)),
    median_divergence_km = round(median(divergence_km, na.rm = TRUE), 1),
    max_divergence_km = round(max(divergence_km, na.rm = TRUE), 1),
    n_reviewed = sum(recommendation != "unreviewed"),
    .groups = "drop"
  ) |>
  arrange(desc(n)) |>
  kable(col.names = c("Presence", "Facilities", "Median peak ADP",
                       "Max peak ADP", "Median div. (km)", "Max div. (km)",
                       "Reviewed"))
```

The **Both FY26 + DDP** group contains the most operationally significant facilities — active detention centers with substantial daily populations. These should be prioritized for geocoding review. The **Neither** group is mostly inactive federal prisons, hold rooms, and medical facilities where divergence is less consequential.

```{r}
#| label: presence-review-status
divergent_classified |>
  count(presence, recommendation) |>
  tidyr::pivot_wider(names_from = recommendation, values_from = n, values_fill = 0) |>
  arrange(desc(unreviewed)) |>
  kable()
```

### High-priority unreviewed facilities

Divergent facilities in the **Both FY26** or **DDP** group that have not yet been assigned a geocode source preference, sorted by peak ADP:

```{r}
#| label: high-priority-unreviewed
divergent_classified |>
  filter(presence %in% c("Both FY26 + DDP", "DDP only", "FY26 only"),
         recommendation == "unreviewed") |>
  arrange(desc(best_peak_adp)) |>
  select(
    `ID` = canonical_id,
    `Facility` = canonical_name,
    `City` = facility_city,
    `State` = facility_state,
    `Source` = source_name,
    `Peak ADP` = best_peak_adp,
    `Divergence (km)` = divergence_km
  ) |>
  kable()
```

## All divergent facilities

```{r}
#| label: divergent-table
divergent_classified |>
  arrange(desc(divergence_km)) |>
  select(
    `ID` = canonical_id,
    `Facility` = canonical_name,
    `Address` = facility_address,
    `City` = facility_city,
    `State` = facility_state,
    `ZIP` = facility_zip,
    `Source` = source_name,
    `Presence` = presence,
    `Rec` = recommendation,
    `Divergence (km)` = divergence_km,
    `Peak ADP` = best_peak_adp,
    `Google lat` = google_lat,
    `Google lon` = google_lon,
    `Source lat` = source_lat,
    `Source lon` = source_lon
  ) |>
  kable()
```
