---
title: "Wikipedia Wikitext Generator"
subtitle: "Copyable {{Infobox prison}} and {{Bar chart}} wikitext for ICE detention facilities"
format:
  html:
    code-fold: true
    toc: true
    toc-depth: 2
---

```{r}
#| label: setup
#| include: false

# Default quarto command to render to quarto-website:
# quarto render infobox-generator.qmd --output-dir ../quarto-website/docs/private/
library(targets)
library(dplyr)
library(htmltools)
source("R/crosswalk.R")
source("R/wikipedia.R")
source("scripts/gh-copy-assets.R")

fw   <- tar_read(facilities_fy26_wiki)
fp   <- tar_read(facility_presence)
cf   <- tar_read(panel_facilities)
geo  <- tar_read(facilities_geocoded_full)
panel <- tar_read(facilities_panel)
xwalk <- tar_read(facility_crosswalk)
```

```{r}
#| label: helpers
#| include: false

wikitext_block <- function(id, label, content) {
  tagList(
    tags$div(
      style = "margin: 8px 0;",
      tags$button(
        class = "copy-btn",
        onclick = sprintf("copyTableToClipboard('%s', this)", id),
        label
      )
    ),
    tags$pre(
      tags$code(content),
      style = "background-color: #f6f8fa; padding: 12px; border-radius: 6px; overflow-x: auto; font-size: 13px;"
    ),
    tags$textarea(
      id = id,
      style = "position: absolute; left: -9999px; height: 0; width: 0; overflow: hidden;",
      readonly = TRUE,
      content
    )
  )
}

wiki_link <- function(slug) {
  if (is.null(slug) || is.na(slug) || slug == "") return(NULL)
  url <- paste0("https://en.wikipedia.org/wiki/", gsub(" ", "_", slug))
  tags$a(href = url, target = "_blank", slug, style = "margin-left: 8px; font-size: 13px;")
}

facility_block <- function(cid, facility_name, facility_city, facility_state,
                           wiki_slug = NA_character_) {
  ib <- generate_infobox(
    cid,
    panel_facilities = cf,
    presence = fp,
    geocoded = geo,
    panel = panel,
    crosswalk = xwalk,
    wiki_matches = fw
  )

  bar <- generate_adp_bar_chart(cid, panel, fp)

  tagList(
    tags$details(
      tags$summary(
        tags$strong(facility_name),
        tags$span(
          paste0(" — ", facility_city, ", ", facility_state),
          style = "color: #586069;"
        ),
        wiki_link(wiki_slug)
      ),
      wikitext_block(paste0("ib_", cid), "Copy Infobox", ib),
      if (nzchar(bar)) wikitext_block(paste0("bar_", cid), "Copy Bar Chart", bar)
    ),
    tags$hr(style = "border: none; border-top: 1px solid #e1e4e8; margin: 6px 0;")
  )
}

render_facility_blocks <- function(df) {
  blocks <- purrr::pmap(
    df |> select(canonical_id, facility_name, facility_city, facility_state,
                 wiki_slug),
    \(canonical_id, facility_name, facility_city, facility_state, wiki_slug)
      facility_block(canonical_id, facility_name, facility_city, facility_state,
                     wiki_slug)
  )
  tagList(blocks)
}
```

```{r}
#| label: classify-facilities
#| include: false

has_article <- fw |>
  filter(!is.na(wiki_slug), wiki_slug != "") |>
  arrange(facility_name)

has_match_only <- fw |>
  filter((is.na(wiki_slug) | wiki_slug == ""),
         !is.na(wiki_match), wiki_match != "") |>
  arrange(facility_name)

redlinks <- fw |>
  filter(is.na(wiki_slug) | wiki_slug == "",
         is.na(wiki_match) | wiki_match == "") |>
  arrange(facility_name)
```

```{r}
#| label: copy-assets
#| echo: false
#| results: asis
include_table_assets()
```

## Existing Wikipedia articles

**`r nrow(has_article)`** facilities with a confirmed Wikipedia article. Use these to add or update infoboxes and population charts on existing pages.

```{r}
#| label: existing-articles
#| echo: false
#| results: asis
render_facility_blocks(has_article)
```

## Unmatched facilities with a fuzzy Wikipedia match

**`r nrow(has_match_only)`** facilities matched to a Wikipedia title by name similarity but not confirmed as the correct article. Check the linked page before using.

```{r}
#| label: fuzzy-matches
#| echo: false
#| results: asis

# Show the fuzzy wiki_match as the link for review
render_facility_blocks(
  has_match_only |> mutate(wiki_slug = wiki_match)
)
```

## Red-link facilities

**`r nrow(redlinks)`** facilities with no Wikipedia article or match. Each section contains wikitext ready to paste into a new article.

```{r}
#| label: redlinks
#| echo: false
#| results: asis
render_facility_blocks(redlinks)
```
