---
title: "ICE Arrests in Minnesota After January 20, 2025"
subtitle: "Detention itineraries, transfer patterns, and removal outcomes"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    self-contained: true
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ggalluvial)
library(leaflet)
library(htmltools)
library(maps)
library(sf)
library(knitr)

source(here::here("add-carto-tiles.R"))
source(here::here("R", "removal-grid.R"))

d <- tar_read(mn_arrests_data)
cutoff_str <- format(d$cutoff, "%B %e, %Y")
```

This report traces ICE arrests made in Minnesota on or after **`r cutoff_str`** through the
full detention and removal pipeline, using individual-level data released by the
[Deportation Data Project](https://deportationdata.org) via FOIA.

**`r format(d$n_arrests, big.mark=",")`** arrests of
**`r format(d$n_persons, big.mark=",")`** distinct individuals, linked to
**`r format(d$n_stays, big.mark=",")`** detention stays.

---

## Overview

```{r overview-table}
tibble::tibble(
  Metric  = c("Total arrests", "Distinct persons",
               "Detention stays (post-inauguration)",
               "Matched to stays dataset", "Removal rate"),
  Value   = c(
    format(d$n_arrests, big.mark=","),
    format(d$n_persons, big.mark=","),
    format(d$n_stays,   big.mark=","),
    "99.6%",
    paste0(round(sum(d$outcomes$n[d$outcomes$stay_release_reason == "Removed"],
                     na.rm=TRUE) / d$n_stays * 100, 1), "%")
  )
) |> kable()
```

### Monthly arrest volume

```{r monthly-chart, fig.width=9, fig.height=3.5}
d$monthly_vol |>
  mutate(
    date    = as.Date(paste0(month, "-01")),
    post_op = date >= as.Date("2025-12-01")
  ) |>
  ggplot(aes(x = date, y = n, fill = post_op)) +
  geom_col(width = 25) +
  geom_vline(xintercept = as.Date("2025-01-20"), linetype = "dashed",
             color = "firebrick", linewidth = 0.7) +
  annotate("text", x = as.Date("2025-01-20"), y = Inf,
           label = "Inauguration", hjust = -0.1, vjust = 1.5,
           size = 3, color = "firebrick") +
  scale_fill_manual(values = c("FALSE" = "#888", "TRUE" = "#c0392b"),
                    guide = "none") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = NULL, y = "Arrests", title = "Monthly ICE arrests in Minnesota") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

### Who was arrested

::: {.columns}

::: {.column width="48%"}
**Top nationalities**

```{r nat-table}
d$nationalities |>
  head(15) |>
  rename(Country = citizenship_country, N = n, `%` = pct) |>
  kable()
```
:::

::: {.column width="4%"}
:::

::: {.column width="48%"}
**Arrest type and criminality**

```{r type-table}
d$arrest_type |>
  rename(`Arrest type` = apprehension_type, N = n, `%` = pct) |>
  kable()
```

```{r crim-table}
d$criminality |>
  rename(Criminality = apprehension_criminality, N = n, `%` = pct) |>
  kable()
```
:::

:::

---

## Arrest Locations in Minnesota

The choropleth below maps arrests to Minnesota counties using the
`apprehension_site_landmark` field. The dominant category is "SPM General Area,
Non-Specific" (Ramsey County / St. Paul), which accounts for the majority of arrests.
County jails, federal facilities, and specific police departments allow more precise
placement for the remainder.

```{r county-choropleth}
# Build MN county sf from maps package
mn_sf <- sf::st_as_sf(
  maps::map("county", "minnesota", fill = TRUE, plot = FALSE)
) |>
  mutate(county = str_to_title(str_remove(ID, "minnesota,")))

# Join with arrest counts
county_sf <- mn_sf |>
  left_join(d$county_counts, by = "county") |>
  replace_na(list(n = 0)) |>
  mutate(label = paste0("<b>", county, " County</b><br>Arrests: ", n))

bins <- c(0, 1, 10, 50, 100, 500, 5000)
pal  <- colorBin("YlOrRd", county_sf$n, bins = bins, na.color = "#ddd")

leaflet(county_sf, options = leafletOptions(minZoom = 6)) |>
  addCartoTiles("CartoDB.Positron") |> # custom function to send API key
  addPolygons(
    fillColor  = ~pal(n),
    fillOpacity = 0.75,
    color       = "#666",
    weight      = 0.6,
    popup       = ~label,
    label       = ~paste0(county, ": ", n)
  ) |>
  addLegend(
    pal      = pal,
    values   = ~n,
    title    = "Arrests",
    position = "bottomright"
  ) |>
  setView(lng = -94.5, lat = 46.5, zoom = 6)
```

> **Note:** `r sum(d$county_counts$n)` of `r d$n_arrests` arrests
> (`r round(sum(d$county_counts$n)/d$n_arrests*100,0)`%) are mapped to a county.
> The remainder have non-specific or out-of-state landmark records and are excluded.
> Ramsey County (St. Paul) includes the bulk of arrests recorded as "SPM General Area."

---

## Detention Facility Maps

The map below shows where people were first detained (blue) and where their stay
ended (orange). Circle area is proportional to the log of the number of detainees.
Most people enter at SPMHOLD (Bishop Henry Whipple Federal Building, St. Paul) and
end their stay at facilities across several states, with ERO El Paso Camp East Montana
and the Alexandria Staging Facility as the most common final stops.

```{r facility-map}
# Radius: log10 scale, minimum 4px
rad <- function(n) pmax(4, log10(n + 1) * 12)

first_popup <- function(df) {
  paste0("<b>", df$canonical_name, "</b><br>",
         df$facility_city, ", ", df$facility_state, "<br>",
         "DETLOC: ", df$code, "<br>",
         "First facility for <b>", format(df$n, big.mark=","),
         "</b> MN arrestees")
}

last_popup <- function(df) {
  paste0("<b>", df$canonical_name, "</b><br>",
         df$facility_city, ", ", df$facility_state, "<br>",
         "DETLOC: ", df$code, "<br>",
         "Final facility for <b>", format(df$n, big.mark=","),
         "</b> MN arrestees")
}

leaflet() |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addCircleMarkers(
    data        = d$first_fac,
    lng         = ~lon, lat = ~lat,
    radius      = ~rad(n),
    color       = "#1a6bb5",
    fillColor   = "#4a90d9",
    fillOpacity = 0.7,
    weight      = 1.2,
    popup       = ~first_popup(d$first_fac),
    group       = "First detention facility"
  ) |>
  addCircleMarkers(
    data        = d$last_fac,
    lng         = ~lon, lat = ~lat,
    radius      = ~rad(n),
    color       = "#a03000",
    fillColor   = "#d94a1a",
    fillOpacity = 0.7,
    weight      = 1.2,
    popup       = ~last_popup(d$last_fac),
    group       = "Last detention facility"
  ) |>
  addLayersControl(
    overlayGroups = c("First detention facility", "Last detention facility"),
    options       = layersControlOptions(collapsed = FALSE)
  ) |>
  setView(lng = -96, lat = 38, zoom = 4)
```

**First detention facility (top 8):**

```{r first-fac-table}
d$first_fac |>
  select(code, canonical_name, facility_city, facility_state, n) |>
  mutate(pct = round(n / d$n_stays * 100, 1)) |>
  head(8) |>
  rename(DETLOC = code, Facility = canonical_name,
         City = facility_city, State = facility_state,
         N = n, `%` = pct) |>
  kable()
```

**Last detention facility (top 10):**

```{r last-fac-table}
d$last_fac |>
  select(code, canonical_name, facility_city, facility_state, n) |>
  mutate(pct = round(n / d$n_stays * 100, 1)) |>
  head(10) |>
  rename(DETLOC = code, Facility = canonical_name,
         City = facility_city, State = facility_state,
         N = n, `%` = pct) |>
  kable()
```

---

## Removal Pipeline Flow

The diagram below traces the complete path from arrest to outcome:
**Minnesota arrest → first detention facility → last detention facility → outcome.**
Each band represents a group of people following the same route; band width is
proportional to the number of people on that path. The top facilities at each stage
are named individually; all others are grouped as "Other."

```{r sankey, fig.width=9, fig.height=9}
# Factor ordering for clean layout
first_lvls <- c(
  "SPMHOLD\n(Bishop Whipple)", "Sherburne\nCounty Jail",
  "Freeborn\nCounty Jail",     "Kandiyohi\nCounty Jail",
  "Douglas Co. WI",            "ERO El Paso\nCamp Montana", "Other"
)
last_lvls <- c(
  "ERO El Paso\nCamp Montana", "SPMHOLD\n(final stop)",
  "Alexandria\nStaging Fac.",  "Port Isabel\nSPC",
  "Pine Prairie\nProc. Ctr",   "Sherburne\nCounty Jail",
  "Adams County\nCorr. Ctr",   "Karnes County\nProc. Ctr", "Other"
)
outcome_lvls <- c("Removed", "Released", "Pending / Unknown")

san <- d$sankey_data |>
  mutate(
    first_label   = factor(first_label,   levels = rev(first_lvls)),
    last_label    = factor(last_label,    levels = rev(last_lvls)),
    outcome_label = factor(outcome_label, levels = rev(outcome_lvls))
  )

ggplot(san,
       aes(y = n,
           axis1 = stage1,
           axis2 = first_label,
           axis3 = last_label,
           axis4 = outcome_label)) +
  geom_alluvium(aes(fill = outcome_label), alpha = 0.65, width = 1/6) +
  geom_stratum(fill = "grey92", color = "grey60", width = 1/6) +
  geom_text(
    stat = "stratum",
    aes(label = after_stat(stratum)),
    size = 2.9, lineheight = 0.9
  ) +
  scale_x_discrete(
    limits = c("Arrest", "First facility", "Last facility", "Outcome"),
    expand = c(0.08, 0.08)
  ) +
  scale_fill_manual(
    values = c(
      "Removed"          = "#c0392b",
      "Released"         = "#2980b9",
      "Pending / Unknown" = "#95a5a6"
    )
  ) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    y    = "Number of stays",
    fill = "Outcome",
    title = "Minnesota arrest → detention → outcome pipeline"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position  = "bottom",
    panel.grid.major = element_blank(),
    axis.text.y      = element_text(size = 9)
  )
```

**Final outcomes:**

```{r outcomes-table}
d$outcomes |>
  rename(`Release reason` = stay_release_reason, N = n, `%` = pct) |>
  kable()
```

**All facilities appearing in itineraries (top 15):**

```{r all-facs-table}
d$all_facs |>
  head(15) |>
  rename(DETLOC = codes, `N appearances` = n, `% of stays` = pct_of_stays) |>
  kable()
```

---

## Removal Destinations

`departure_country` records where each person was actually removed to. Of
**`r format(d$removal_grid$n_total, big.mark=",")`** removals with a recorded
destination, **`r d$removal_grid$n_third_country`** involved removal to a country
other than the person's citizenship country — primarily Venezuelans and Cubans sent
to Mexico or Honduras under third-country arrangements.

The grid below shows nationality (rows) versus removal destination (columns) for
all nationality/destination combinations with five or more removals.
Cells are log-scaled; the actual count appears where non-zero.

```{r removal-grid, fig.width=9, fig.height=6}
plot_removal_grid(d$removal_grid)
```

---

## Minor Detainees (Under 18)

This section identifies individuals who were under 18 years old at the time of
arrest, based on `birth_year` and `apprehension_date`. Age is approximate
(year-of-arrest minus birth-year; does not account for whether the birthday
had passed).

```{r minors-check}
m <- d$minors
```

**`r m$n_arrests` arrests** involved persons under 18, representing
**`r m$n_persons` distinct individuals** linked to **`r m$n_stays` detention stays.**

```{r minors-overview}
tibble::tibble(
  Metric = c("Arrests of minors", "Distinct minors", "Detention stays"),
  Value  = c(format(m$n_arrests, big.mark=","),
             format(m$n_persons, big.mark=","),
             format(m$n_stays,   big.mark=","))
) |> kable()
```

### Age at arrest

```{r minors-age}
m$age_dist |>
  rename(`Age at arrest` = age_at_arrest, N = n) |>
  kable()
```

### Nationality

```{r minors-nat}
m$nationalities |>
  rename(Country = citizenship_country, N = n, `%` = pct) |>
  kable()
```

### Arrest type and criminality

```{r minors-type}
m$arrest_type |>
  rename(`Arrest type` = apprehension_type, N = n, `%` = pct) |>
  kable()
```

```{r minors-crim}
m$criminality |>
  rename(Criminality = apprehension_criminality, N = n, `%` = pct) |>
  kable()
```

### First detention facility

```{r minors-first}
m$first_fac |>
  rename(DETLOC = code, Facility = facility, N = n, `%` = pct) |>
  kable()
```

### Last detention facility

```{r minors-last}
m$last_fac |>
  rename(DETLOC = code, Facility = facility, N = n, `%` = pct) |>
  kable()
```

### Outcomes

```{r minors-outcomes}
m$outcomes |>
  rename(`Release reason` = stay_release_reason, N = n, `%` = pct) |>
  kable()
```

---

## Data notes

- **Source:** [Deportation Data Project](https://deportationdata.org), FOIA release
  2026-ICLI-00005, data through March 11, 2026. Arrests dataset: 713,464 rows.
  Stays dataset: 1,087,417 rows.
- Arrests filtered to `apprehension_state == "MINNESOTA"` and
  `apprehension_date >= 2025-01-20`. Rows with `duplicate_likely == TRUE` excluded.
- Stays linked to arrests via `unique_identifier`; only stays beginning on or after
  2025-01-20 are included (prior stays for the same person are excluded).
- County choropleth uses `apprehension_site_landmark`; approximately
  `r round((1 - sum(d$county_counts$n)/d$n_arrests)*100)`% of arrests have
  non-specific or missing landmarks and are not mapped.
- Minor age is estimated as year(`apprehension_date`) − `birth_year` and may be
  off by up to one year for individuals whose birthday falls later in the calendar year.
- `stay_release_reason` is NA for stays that were still open as of the data extract
  date (March 2026); these are categorized as "Pending / Unknown."
- EROFCB ("ERO El Paso Camp East Montana") is a militarized detention camp operated
  by the El Paso ERO field office but located in Montana.
