---
title: "ICE Arrests Nationwide: Geographic Distribution"
subtitle: "Post-inauguration arrests from the DDP FOIA dataset"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-float: true
    self-contained: true
params:
  cutoff: "2025-01-20"
execute:
  echo: false
  warning: false
  message: false
---

Notes for next improvements: 
1. Include Alaska and Hawaii on the national maps.
2. The county-level data excludes a lot of towns, Louisiana's parishes, 
and many metropolitan areas that don't follow the "NAME COUNTY" pattern.
(We should examine the share of arrests in each state that make 
each of the patterns.)
3. We could create an interactive line graph with all of the state arrest
rates.

```{r setup}
library(targets)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(leaflet)
library(sf)
library(maps)
library(knitr)

cutoff     <- as.Date(params$cutoff)
cutoff_str <- format(cutoff, "%B %e, %Y")
data_end   <- as.Date("2026-03-31")   # last full month of DDP data; 2026-09 is a data error

arrests_raw <- tar_read(arrests_raw)
stays_raw   <- tar_read(stays_raw)
source(here::here("R/state-arrests.R"))        # provides STATE_ABBREVS
source(here::here("R/state-arrests-export.R")) # provides build_state_outcomes()
source(here::here("R/removal-grid.R"))         # provides build_removal_grid(), plot_removal_grid()

# Full deduplicated dataset — used for the time-series trend chart
arr_all <- arrests_raw |>
  filter(duplicate_likely == FALSE | is.na(duplicate_likely))

# Post-cutoff only — used for maps, tables, and counts
arr <- arr_all |>
  filter(apprehension_date >= cutoff)

n_total      <- nrow(arr)
n_with_state <- sum(!is.na(arr$apprehension_state))
```

Maps and tables below cover arrests on or after **`r cutoff_str`**.
**`r format(n_total, big.mark=",")`** arrests ·
**`r round(n_with_state/n_total*100,1)`%** have a state recorded.
The monthly trend chart shows the full dataset for context.

---

## State Bubble Map

Circle **area** is proportional to the number of arrests.
Circle **color** shows arrests per 1,000 estimated undocumented immigrants
(Pew 2023 estimates), a rough intensity measure independent of state size.
Popups include Census 2025 total population and per-capita rates.

```{r state-counts}
state_counts <- arr |>
  filter(!is.na(apprehension_state)) |>
  mutate(state_lower = str_to_lower(apprehension_state)) |>
  count(state_lower, sort = TRUE)

# Census 2025 population estimates (column 8 = July 1 2025)
census_pops <- readxl::read_xlsx(
  "data/census/NST-EST2025-POP.xlsx", col_names = FALSE
) |>
  select(geo_area = 1, pop_2025 = 8) |>
  filter(!is.na(geo_area), str_starts(geo_area, "\\.")) |>
  mutate(
    state_lower = str_to_lower(str_remove(geo_area, "^\\.")),
    pop_2025    = as.numeric(pop_2025)
  ) |>
  select(state_lower, pop_2025)

# Pew 2023 undocumented immigrant estimates
pew_undoc <- readxl::read_xlsx(
  "data/pew/RE_2025.08.21_Unauthorized-immigrants_detailed-tables_characteristics-for-states.xlsx",
  sheet = "Population Characteristics", col_names = FALSE, skip = 4
) |>
  select(state_name = 2, undoc_pop = 3, undoc_pct_pop = 4) |>
  filter(!is.na(state_name), state_name != "U.S., total",
         !str_starts(state_name, "Note")) |>
  mutate(
    state_lower   = str_to_lower(state_name),
    undoc_pop     = as.numeric(undoc_pop),
    undoc_pct_pop = as.numeric(undoc_pct_pop)
  ) |>
  select(state_lower, undoc_pop, undoc_pct_pop)

# Join arrests + Census + Pew
state_data <- state_counts |>
  rename(n_arrests = n) |>
  left_join(census_pops, by = "state_lower") |>
  left_join(pew_undoc,   by = "state_lower") |>
  mutate(
    state_title          = str_to_title(state_lower),
    arrests_per_1k_undoc = if_else(
      !is.na(undoc_pop) & undoc_pop > 0,
      round(n_arrests / undoc_pop * 1000, 2), NA_real_
    ),
    arrests_per_100k_pop = if_else(
      !is.na(pop_2025) & pop_2025 > 0,
      round(n_arrests / pop_2025 * 1e5, 1), NA_real_
    )
  )
```

```{r state-map}
us_states_sf <- sf::st_as_sf(
  maps::map("state", fill = TRUE, plot = FALSE)
) |>
  rename(state_lower = ID)

# State centroids — disable s2 to tolerate minor self-intersections from maps pkg
sf::sf_use_s2(FALSE)
cent_sf    <- sf::st_centroid(us_states_sf)
cent_coords <- sf::st_coordinates(cent_sf)
centroids  <- cent_sf |>
  sf::st_drop_geometry() |>
  mutate(lon = cent_coords[, 1], lat = cent_coords[, 2])
sf::sf_use_s2(TRUE)

bubble_data <- centroids |>
  left_join(state_data, by = "state_lower") |>
  filter(!is.na(n_arrests), n_arrests > 0)

# Color: arrests per 1k undoc
pal_bubble <- colorNumeric(
  palette  = c("#fff5f0", "#fc8d59", "#a50f15"),
  domain   = bubble_data$arrests_per_1k_undoc,
  na.color = "#bbbbbb"
)

# Radius: log10-scaled, in pixels
rad_state <- function(n) pmax(5, sqrt(n + 1) / 5)

# Rich popups
bubble_data <- bubble_data |>
  mutate(
    state_abbr = unname(STATE_ABBREVS[str_to_upper(state_lower)]),
    popup = paste0(
      "<b style='font-size:1.1em'>", state_title, "</b>",
      if_else(!is.na(state_abbr),
        paste0(" <a href='arrests-by-state/", state_abbr, ".html'",
               " style='font-size:0.85em'>details →</a>"), ""),
      "<hr style='margin:3px 0'>",
      "Post-inauguration arrests: <b>", format(n_arrests, big.mark = ","), "</b><br>",
      if_else(!is.na(arrests_per_1k_undoc),
        paste0("Arrests / 1,000 undocumented: <b>",
               arrests_per_1k_undoc, "</b><br>"), ""),
      if_else(!is.na(undoc_pop),
        paste0("Est. undocumented pop. (Pew 2023): ",
               format(undoc_pop, big.mark = ","),
               " (", round(undoc_pct_pop * 100, 1), "% of state)<br>"), ""),
      if_else(!is.na(pop_2025),
        paste0("State population (Census 2025): ",
               format(round(pop_2025), big.mark = ","), "<br>"), ""),
      if_else(!is.na(arrests_per_100k_pop),
        paste0("Arrests / 100k total population: ", arrests_per_100k_pop), "")
    )
  )

leaflet(bubble_data) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addCircleMarkers(
    lng         = ~lon, lat = ~lat,
    radius      = ~rad_state(n_arrests),
    fillColor   = ~pal_bubble(arrests_per_1k_undoc),
    fillOpacity = 0.75,
    color       = "#444",
    weight      = 0.7,
    popup       = ~popup,
    label       = ~paste0(state_title, ": ",
                           format(n_arrests, big.mark = ","), " arrests; ",
                           arrests_per_1k_undoc, "/1k undoc.")
  ) |>
  addLegend(
    pal      = pal_bubble,
    values   = ~arrests_per_1k_undoc,
    title    = "Arrests per<br>1,000 undoc.",
    position = "bottomright",
    labFormat = labelFormat(suffix = "")
  ) |>
  setView(lng = -96, lat = 38, zoom = 4)
```

**Top 20 states by arrests:**
```{r state-table}
state_data |>
  select(state_title, n_arrests, arrests_per_1k_undoc, undoc_pop, pop_2025) |>
  head(20) |>
  mutate(
    pct_of_arrests = round(n_arrests / sum(state_data$n_arrests) * 100, 1),
    undoc_pop      = format(undoc_pop, big.mark = ",", scientific = FALSE),
    pop_2025       = format(round(pop_2025), big.mark = ",", scientific = FALSE)
  ) |>
  rename(
    State                   = state_title,
    Arrests                 = n_arrests,
    `% of arrests`          = pct_of_arrests,
    `Arrests/1k undoc.`     = arrests_per_1k_undoc,
    `Undoc. pop. (Pew 2023)` = undoc_pop,
    `Population (2025)`      = pop_2025
  ) |>
  kable()
```

---

## County-Level Choropleth

County names are extracted from `apprehension_site_landmark` using a
"[NAME] COUNTY" pattern.

```{r county-counts}
county_counts <- arr |>
  filter(!is.na(apprehension_site_landmark),
         !is.na(apprehension_state)) |>
  mutate(
    state_lower = str_to_lower(apprehension_state),
    county_raw  = str_remove(
      str_extract(str_to_upper(apprehension_site_landmark),
                  "^([A-Z .'-]+) COUNTY"),
      " COUNTY$"
    ),
    county_lower = str_to_lower(county_raw),
    map_id = paste0(state_lower, ",", county_lower)
  ) |>
  filter(!is.na(county_lower)) |>
  count(map_id, sort = TRUE)

county_pct <- round(sum(county_counts$n) / n_total * 100, 1)
```

> **Coverage:** `r county_pct`% of post-`r cutoff_str` arrests mapped to a county.

```{r county-map}
us_counties_sf <- sf::st_as_sf(
  maps::map("county", fill = TRUE, plot = FALSE)
) |>
  rename(map_id = ID)

county_sf <- us_counties_sf |>
  left_join(county_counts, by = "map_id") |>
  replace_na(list(n = 0)) |>
  mutate(
    county_title = str_to_title(str_extract(map_id, "(?<=,).*")),
    state_title  = str_to_title(str_extract(map_id, "^[^,]+")),
    # NA for 0-1 arrests → rendered transparent
    n_map = if_else(n <= 1, NA_real_, as.double(n)),
    label = paste0("<b>", county_title, " County, ", state_title,
                   "</b><br>Arrests: ", format(n, big.mark=","))
  )

bins_c <- c(1, 5, 20, 100, 500, max(county_sf$n, na.rm = TRUE) + 1)
pal_c  <- colorBin("YlOrRd", county_sf$n_map, bins = bins_c, na.color = "transparent")

leaflet(county_sf) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(
    fillColor   = ~pal_c(n_map), fillOpacity = 0.7,
    color = "#ccc", weight = 0.15,
    popup = ~label,
    label = ~if_else(n > 0, paste0(county_title, ", ", state_title, ": ", n), NA_character_)
  ) |>
  addLegend(pal = pal_c, values = ~n_map, title = "Arrests",
            position = "bottomright") |>
  setView(lng = -96, lat = 38, zoom = 4)
```

> **Note:** Counties with zero arrests shown in light grey may have arrests recorded
> under non-specific landmarks. Alaska and Hawaii are excluded from the `maps` package
> continental US boundary data.

**Top 30 counties by arrests:**
```{r county-table}
county_counts |>
  mutate(
    county_title = str_to_title(str_extract(map_id, "(?<=,).*")),
    state_title  = str_to_title(str_extract(map_id, "^[^,]+")),
    pct = round(n / sum(n) * 100, 1)
  ) |>
  select(county_title, state_title, n, pct) |>
  head(30) |>
  rename(County = county_title, State = state_title, N = n, `%` = pct) |>
  kable()
```

---

## Monthly Trends by Top State

```{r monthly-trends, fig.width=11, fig.height=6}
# Use full dataset (capped at data_end) for trend context
# Top states ranked by post-cutoff arrest volume
top_states <- state_counts |> head(8) |> pull(state_lower)

trend_base <- arr_all |>
  filter(!is.na(apprehension_state), apprehension_date <= data_end) |>
  mutate(
    state_lower = str_to_lower(apprehension_state),
    month       = as.Date(paste0(format(apprehension_date, "%Y-%m"), "-01"))
  )

trend_base |>
  mutate(state_grp = if_else(state_lower %in% top_states,
                              str_to_title(state_lower), "Other")) |>
  count(state_grp, month) |>
  filter(state_grp != "Other") |>
  ggplot(aes(x = month, y = n, color = state_grp)) +
  geom_line(linewidth = 0.9) +
  geom_vline(xintercept = cutoff, linetype = "dashed",
             color = "firebrick", linewidth = 0.6) +
  annotate("text", x = cutoff, y = Inf,
           label = "Inauguration", hjust = -0.1, vjust = 1.5,
           size = 3, color = "firebrick") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months",
               limits = c(min(trend_base$month), data_end)) +
  scale_y_continuous(labels = scales::comma) +
  scale_color_brewer(palette = "Set2") +
  labs(x = NULL, y = "Arrests", color = "State",
       title = "Monthly ICE arrests by state — top 8 by post-inauguration volume") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "right")
```

## Arrest Rate Change by State

For each state, compute arrests **per day** in the before and after periods, then
take the ratio. Values above 1 mean more arrests per day after inauguration;
values below 1 mean fewer. Only states with at least 30 arrests in the before
period are included.

```{r ratio-calc}
before_days <- as.numeric(cutoff        - as.Date("2022-10-01"))
after_days  <- as.numeric(data_end + 1  - cutoff)

state_ratio <- arr_all |>
  filter(!is.na(apprehension_state), apprehension_date <= data_end) |>
  mutate(
    state_lower = str_to_lower(apprehension_state),
    period      = if_else(apprehension_date < cutoff, "before", "after")
  ) |>
  count(state_lower, period) |>
  tidyr::pivot_wider(names_from = period, values_from = n, values_fill = 0L) |>
  mutate(
    rate_before    = before / before_days,
    rate_after     = after  / after_days,
    ratio          = if_else(rate_before > 0, round(rate_after / rate_before, 2), NA_real_),
    state_title    = str_to_title(state_lower)
  ) |>
  filter(!is.na(ratio), before >= 30) |>
  arrange(desc(ratio))
```

```{r state-outcomes}
# Removal outcomes per state (single-pass over stays)
state_outcomes_tbl <- build_state_outcomes(arr, stays_raw, cutoff)
```

**All states and DC — arrest rate change and removal outcomes:**
```{r ratio-reactable}
library(reactable)

# Join ratio + outcomes + Census/Pew + abbrev for links
reactable_data <- state_ratio |>
  filter(before >= 5) |>
  left_join(state_data |> select(state_lower, n_arrests, arrests_per_1k_undoc,
                                   undoc_pop, pop_2025),
            by = "state_lower") |>
  left_join(state_outcomes_tbl, by = "state_lower") |>
  mutate(
    state_title = str_to_title(state_lower),
    state_abbr  = unname(STATE_ABBREVS[str_to_upper(state_lower)]),
    rate_before = round(rate_before, 2),
    rate_after  = round(rate_after,  2)
  ) |>
  filter(!is.na(state_abbr)) |>
  arrange(desc(ratio))

reactable(
  reactable_data,
  defaultPageSize = 15,
  searchable      = TRUE,
  striped         = TRUE,
  highlight       = TRUE,
  columns = list(
    state_lower      = colDef(show = FALSE),
    state_abbr       = colDef(show = FALSE),
    before           = colDef(show = FALSE),
    after            = colDef(show = FALSE),
    pop_2025         = colDef(show = FALSE),
    n_stays_outcomes = colDef(show = FALSE),
    removed          = colDef(show = FALSE),
    released         = colDef(show = FALSE),
    pending          = colDef(show = FALSE),
    state_title = colDef(
      name = "State", minWidth = 130,
      cell = function(value, index) {
        abbr <- reactable_data$state_abbr[index]
        if (!is.na(abbr)) {
          htmltools::tags$a(
            href = paste0("arrests-by-state/", abbr, ".html"),
            value
          )
        } else value
      }
    ),
    n_arrests           = colDef(name = "Arrests (after)", format = colFormat(separators = TRUE)),
    rate_before         = colDef(name = "Per day (before)", format = colFormat(digits = 2)),
    rate_after          = colDef(name = "Per day (after)",  format = colFormat(digits = 2)),
    ratio               = colDef(
      name = "Ratio", format = colFormat(digits = 2),
      style = function(value) {
        if (is.na(value)) return(list())
        bg <- if (value >= 5) "#a50f15" else if (value >= 2) "#fc8d59" else if (value >= 1) "#fee08b" else "#d9f0a3"
        list(background = bg, fontWeight = "bold")
      }
    ),
    pct_removed  = colDef(name = "% Removed",  format = colFormat(suffix = "%", digits = 1)),
    pct_released = colDef(name = "% Released", format = colFormat(suffix = "%", digits = 1)),
    pct_pending  = colDef(name = "% Pending",  format = colFormat(suffix = "%", digits = 1)),
    arrests_per_1k_undoc = colDef(
      name = "Arrests/1k undoc.", format = colFormat(digits = 2)
    ),
    undoc_pop = colDef(
      name = "Undoc. pop. (Pew)", format = colFormat(separators = TRUE)
    )
  )
)
```

```{r ratio-trend, fig.width=11, fig.height=6}
top_ratio_states <- state_ratio |> head(10) |> pull(state_lower)

trend_base |>
  filter(state_lower %in% top_ratio_states) |>
  count(state_grp = str_to_title(state_lower), month) |>
  ggplot(aes(x = month, y = n, color = state_grp)) +
  geom_line(linewidth = 0.9) +
  geom_vline(xintercept = cutoff, linetype = "dashed",
             color = "firebrick", linewidth = 0.6) +
  annotate("text", x = cutoff, y = Inf,
           label = "Inauguration", hjust = -0.1, vjust = 1.5,
           size = 3, color = "firebrick") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "3 months",
               limits = c(min(trend_base$month), data_end)) +
  scale_y_continuous(labels = scales::comma) +
  scale_color_brewer(palette = "Paired") +
  labs(x = NULL, y = "Arrests", color = "State",
       title = "Monthly ICE arrests — top 10 states by post/pre arrest rate ratio") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1),
        legend.position = "right")
```

---

## Removal Destinations

```{r nationwide-removal-grid-data}
stays_post <- stays_raw |>
  dplyr::filter(as.Date(stay_book_in_date_time) >= cutoff)

nationwide_rg <- build_removal_grid(stays_post, "Nationwide", min_n = 500)
```

`departure_country` records where each person was actually sent.
Of **`r format(nationwide_rg$n_total, big.mark=",")`** removals with a recorded
destination, **`r format(nationwide_rg$n_third_country, big.mark=",")`**
(`r round(nationwide_rg$n_third_country / nationwide_rg$n_total * 100, 1)`%) involved
removal to a country other than the person's citizenship country.
The grid below shows nationalities and destinations with at least 500 removals.

```{r nationwide-removal-grid, fig.width=10, fig.height=7}
plot_removal_grid(nationwide_rg)
```

---

## Data notes

- Source: [Deportation Data Project](https://deportationdata.org), FOIA 2026-ICLI-00005.
- Maps and tables filtered to arrests on or after **`r cutoff_str`**
  (`r format(n_total, big.mark=",")` arrests). Trend chart shows the full dataset
  from `r format(min(arr_all$apprehension_date), "%B %Y")` –
  `r format(max(arr_all$apprehension_date), "%B %Y")`.
- `apprehension_state` is missing for `r format(sum(is.na(arr$apprehension_state)), big.mark=",")`
  post-cutoff rows (`r round(mean(is.na(arr$apprehension_state))*100,1)`%); excluded
  from state and county counts.
- County extraction uses the regex `^([A-Z .'-]+) COUNTY` on the
  `apprehension_site_landmark` field. Multi-word county names (e.g., "Crow Wing")
  are captured; names embedded mid-string are not.
- Alaska and Hawaii are not shown in the continental US maps.
- **Census population** from U.S. Census Bureau, "Annual Estimates of the Resident
  Population for the United States, Regions, States, District of Columbia and Puerto Rico"
  (NST-EST2025-POP.xlsx), July 1 2025 estimates.
- **Undocumented population** from Pew Research Center, "Unauthorized immigrants: 
  Characteristics for states" (2023 estimates, released 2025). Estimates are rounded
  to the nearest 5,000–50,000 depending on state size.
- Arrest intensity (arrests per 1,000 undocumented) uses Pew 2023 undocumented
  population as denominator; this is the most recent state-level estimate available.
