---
title: "ICE Arrests: `r stringr::str_to_title(stringr::str_to_lower(params$state))` After January 20, 2025"
subtitle: "Detention itineraries, transfer flows, and removal outcomes"
date: today
format:
  html:
    toc: true
    toc-depth: 3
    toc-float: true
    self-contained: true
params:
  state: "ILLINOIS"
  cutoff: "2025-01-20"
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
library(targets)
library(dplyr)
library(tidyr)
library(purrr)
library(stringr)
library(ggplot2)
library(ggalluvial)
library(leaflet)
library(sf)
library(maps)
library(knitr)

source(here::here("R/state-arrests.R"))

# Read raw pipeline targets — works for any state without rebuilding pipeline
arrests_raw   <- tar_read(arrests_raw)
stays_raw     <- tar_read(stays_raw)
detloc_lkp    <- tar_read(detloc_lookup_complete)
geo_all       <- readr::read_csv(here::here("data/facilities-geocoded-all.csv"),
                                  show_col_types = FALSE)

d <- build_state_arrests_data(
  arrests      = arrests_raw,
  stays        = stays_raw,
  detloc_lookup = detloc_lkp,
  geo_all      = geo_all,
  state        = params$state,
  cutoff       = as.Date(params$cutoff)
)

state_full  <- str_to_title(str_to_lower(d$state))
state_lower <- str_to_lower(d$state)
cutoff_str  <- format(d$cutoff, "%B %e, %Y")

removal_n   <- sum(d$outcomes$n[d$outcomes$stay_release_reason == "Removed"],
                   na.rm = TRUE)
removal_pct <- round(removal_n / d$n_stays * 100, 1)
```

This report traces ICE arrests made in **`r state_full`** on or after **`r cutoff_str`**
through the full detention and removal pipeline, using individual-level FOIA data
from the [Deportation Data Project](https://deportationdata.org).

**`r format(d$n_arrests, big.mark=",")`** arrests ·
**`r format(d$n_persons, big.mark=",")`** distinct individuals ·
**`r format(d$n_stays, big.mark=",")`** detention stays ·
**`r removal_pct`%** ultimately removed

---

## Overview

### Monthly arrest volume

```{r monthly-chart, fig.width=9, fig.height=3.5}
d$monthly_vol |>
  mutate(
    date     = as.Date(paste0(month, "-01")),
    elevated = {
      q75 <- quantile(d$monthly_vol$n, 0.75)
      iqr <- IQR(d$monthly_vol$n)
      n > q75 + 1.5 * iqr
    }
  ) |>
  ggplot(aes(x = date, y = n, fill = elevated)) +
  geom_col(width = 25) +
  geom_vline(xintercept = as.Date(params$cutoff),
             linetype = "dashed", color = "firebrick", linewidth = 0.7) +
  annotate("text", x = as.Date(params$cutoff), y = Inf,
           label = "Inauguration", hjust = -0.1, vjust = 1.5,
           size = 3, color = "firebrick") +
  scale_fill_manual(values = c("FALSE" = "#888", "TRUE" = "#c0392b"),
                    guide = "none") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "2 months") +
  scale_y_continuous(labels = scales::comma) +
  labs(x = NULL, y = "Arrests",
       title = paste("Monthly ICE arrests in", state_full)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

### Who was arrested

:::{.grid}
:::{.g-col-6}
**Top nationalities**
```{r nat-table}
d$nationalities |> head(15) |>
  rename(Country = citizenship_country, N = n, `%` = pct) |> kable()
```
:::
:::{.g-col-6}
**Arrest type**
```{r type-table}
d$arrest_type |>
  rename(`Arrest type` = apprehension_type, N = n, `%` = pct) |> kable()
```
**Criminality at arrest**
```{r crim-table}
d$criminality |>
  rename(Criminality = apprehension_criminality, N = n, `%` = pct) |> kable()
```
:::
:::

---

## Arrest Locations in `r state_full`

County-level map based on `apprehension_site_landmark`. Records with non-specific
landmarks (e.g. "SPM General Area") are mapped using state-specific rules where
possible; others are excluded.

```{r county-choropleth}
county_coverage <- sum(d$county_counts$n) / d$n_arrests

state_map_sf <- tryCatch({
  sf::st_as_sf(
    maps::map("county", state_lower, fill = TRUE, plot = FALSE)
  ) |>
    mutate(county = str_to_title(str_remove(ID, paste0(state_lower, ","))))
}, error = function(e) NULL)

if (!is.null(state_map_sf) && nrow(d$county_counts) > 0 && county_coverage >= 0.20) {
  county_sf <- state_map_sf |>
    left_join(d$county_counts, by = "county") |>
    replace_na(list(n = 0)) |>
    mutate(
      n_map = if_else(n <= 1, NA_real_, as.double(n)),
      label = paste0("<b>", county, " County</b><br>Arrests: ", n)
    )

  max_n <- max(county_sf$n, na.rm = TRUE)
  bins  <- unique(c(1, 5, 20, 100, 500, max_n + 1)[c(1, 5, 20, 100, 500, max_n + 1) <= max_n + 1])
  pal   <- colorBin("YlOrRd", county_sf$n_map, bins = bins, na.color = "transparent")

  leaflet(county_sf) |>
    addCartoTiles("CartoDB.Positron") |> # custom function to send API key
    addPolygons(
      fillColor   = ~pal(n_map), fillOpacity = 0.75,
      color = "#999", weight = 0.4,
      popup = ~label, label = ~paste0(county, ": ", n)
    ) |>
    addLegend(pal = pal, values = ~n_map, title = "Arrests",
              position = "bottomright")
} else {
  # Low county coverage — show top apprehension landmarks instead
  knitr::kable(
    d$landmark_counts |>
      head(30) |>
      rename(`Arrest location` = apprehension_site_landmark, N = n, `%` = pct),
    caption = paste0(
      "Top arrest locations (county data available for only ",
      round(county_coverage * 100, 0), "% of arrests)"
    )
  )
}
```

> **Coverage:** `r round(county_coverage * 100, 0)`% of arrests mapped to a county.
> `r if (county_coverage < 0.20) "Landmark table shown in place of choropleth."`

---

## Detention Facility Maps and Transfer Flows

**Blue dots** = first detention facility. **Orange dots** = last detention facility.
**Grey lines** = transfer flows between facilities; line width is proportional to
the log of the number of transfers on that route. Only flows with ≥
`r max(5L, as.integer(d$n_stays * 0.001))` transfers are shown.

```{r flow-map}
# Build sf linestrings from flow arrows tibble
flows <- d$flow_arrows

rad <- function(n) pmax(3, log10(n + 1) * 8.4)   # ~70% of prior scale

base_map <- leaflet() |>
  addProviderTiles(providers$CartoDB.Positron)

# Geodetic bearing in degrees (0 = North, clockwise)
.bearing <- function(lon1, lat1, lon2, lat2) {
  dlon  <- (lon2 - lon1) * pi / 180
  lat1r <- lat1 * pi / 180;  lat2r <- lat2 * pi / 180
  y <- sin(dlon) * cos(lat2r)
  x <- cos(lat1r) * sin(lat2r) - sin(lat1r) * cos(lat2r) * cos(dlon)
  (atan2(y, x) * 180 / pi + 360) %% 360
}

# SVG arrowhead icon, pointing in bearing direction
.arrow_icon <- function(bearing_deg, size = 10) {
  svg <- paste0(
    '<svg xmlns="http://www.w3.org/2000/svg" viewBox="0 0 12 12"',
    ' width="', size, '" height="', size, '">',
    '<polygon points="6,0 11,11 6,8 1,11"',
    ' fill="#333" opacity="0.75"',
    ' transform="rotate(', round(bearing_deg, 1), ',6,6)"/></svg>'
  )
  makeIcon(
    iconUrl     = paste0("data:image/svg+xml;charset=utf-8,",
                          URLencode(svg, reserved = FALSE)),
    iconWidth   = size, iconHeight = size,
    iconAnchorX = size / 2, iconAnchorY = size / 2
  )
}

# Add flow lines and arrowheads if we have any
if (nrow(flows) > 0) {
  lines_geom <- lapply(seq_len(nrow(flows)), function(i) {
    st_linestring(matrix(
      c(flows$from_lon[i], flows$from_lat[i],
        flows$to_lon[i],   flows$to_lat[i]),
      ncol = 2, byrow = TRUE
    ))
  })
  flows_sf <- st_sf(flows, geometry = st_sfc(lines_geom, crs = 4326))

  base_map <- base_map |>
    addPolylines(
      data    = flows_sf,
      weight  = ~weight,
      color   = "#555",
      opacity = 0.5,
      popup   = ~label,
      group   = "Transfer flows"
    )

}

first_popup <- paste0(
  "<b>", d$first_fac$canonical_name, "</b><br>",
  d$first_fac$facility_city, ", ", d$first_fac$facility_state, "<br>",
  "DETLOC: ", d$first_fac$code, "<br>",
  "First facility for <b>", format(d$first_fac$n, big.mark = ","),
  "</b> arrests"
)
last_popup <- paste0(
  "<b>", d$last_fac$canonical_name, "</b><br>",
  d$last_fac$facility_city, ", ", d$last_fac$facility_state, "<br>",
  "DETLOC: ", d$last_fac$code, "<br>",
  "Final facility for <b>", format(d$last_fac$n, big.mark = ","),
  "</b> arrests"
)

base_map |>
  addCircleMarkers(
    data = d$first_fac, lng = ~lon, lat = ~lat,
    radius = ~rad(n), color = "#1a6bb5", fillColor = "#4a90d9",
    fillOpacity = 0.75, weight = 1.2,
    popup = first_popup, group = "First detention facility"
  ) |>
  addCircleMarkers(
    data = d$last_fac, lng = ~lon, lat = ~lat,
    radius = ~rad(n), color = "#a03000", fillColor = "#d94a1a",
    fillOpacity = 0.75, weight = 1.2,
    popup = last_popup, group = "Last detention facility"
  ) |>
  addLayersControl(
    overlayGroups = c("Transfer flows",
                      "First detention facility",
                      "Last detention facility"),
    options = layersControlOptions(collapsed = FALSE)
  ) |>
  setView(lng = -96, lat = 38, zoom = 4) |>
  hideGroup("Transfer flows")
```

**First detention facility (top 8):**
```{r first-fac-table}
d$first_fac |>
  mutate(pct = round(n / d$n_stays * 100, 1)) |>
  select(code, canonical_name, facility_city, facility_state, n, pct) |>
  head(8) |>
  rename(DETLOC = code, Facility = canonical_name,
         City = facility_city, State = facility_state, N = n, `%` = pct) |>
  kable()
```

**Last detention facility (top 10):**
```{r last-fac-table}
d$last_fac |>
  mutate(pct = round(n / d$n_stays * 100, 1)) |>
  select(code, canonical_name, facility_city, facility_state, n, pct) |>
  head(10) |>
  rename(DETLOC = code, Facility = canonical_name,
         City = facility_city, State = facility_state, N = n, `%` = pct) |>
  kable()
```

---

## Removal Pipeline Flow

Each band traces one group through the pipeline:
**`r state_full` arrest → first facility → last facility → outcome.**
Band width is proportional to number of stays.

```{r sankey, fig.width=14, fig.height=9}
# Build label lookup for display names
geo_lookup_names <- bind_rows(
  d$first_fac |> select(code, name = canonical_name),
  d$last_fac  |> select(code, name = canonical_name)
) |> distinct(code, .keep_all = TRUE)

short_name <- function(code) {
  nm <- geo_lookup_names$name[match(code, geo_lookup_names$code)]
  ifelse(is.na(nm), code,
         paste0(code, "\n(", str_trunc(nm, 22, ellipsis = "…"), ")"))
}

san <- d$sankey_data |>
  mutate(
    first_disp   = if_else(first_label == "Other", "Other",
                           short_name(first_label)),
    last_disp    = if_else(last_label  == "Other", "Other",
                           short_name(last_label)),
    outcome_fct  = factor(outcome_label,
                          levels = c("Removed", "Released", "Pending / Unknown"))
  )

ggplot(san,
       aes(y = n,
           axis1 = stage1, axis2 = first_disp,
           axis3 = last_disp, axis4 = outcome_fct)) +
  geom_alluvium(aes(fill = outcome_fct), alpha = 0.65, width = 1/6) +
  geom_stratum(fill = "grey92", color = "grey60", width = 1/6) +
  geom_text(stat = "stratum",
            aes(label = after_stat(stratum)),
            size = 2.8, lineheight = 0.9) +
  scale_x_discrete(
    limits = c("Arrest", "First facility", "Last facility", "Outcome"),
    expand = c(0.08, 0.08)
  ) +
  scale_fill_manual(values = c(
    "Removed"           = "#c0392b",
    "Released"          = "#2980b9",
    "Pending / Unknown" = "#95a5a6"
  )) +
  scale_y_continuous(labels = scales::comma) +
  labs(y = "Number of stays", fill = "Outcome",
       title = paste(state_full, "arrest → detention → outcome pipeline")) +
  theme_minimal(base_size = 12) +
  theme(legend.position = "bottom", panel.grid.major = element_blank(),
        axis.text.y = element_text(size = 9))
```

**Final outcomes:**
```{r outcomes-table}
d$outcomes |>
  rename(`Release reason` = stay_release_reason, N = n, `%` = pct) |> kable()
```

**All facilities in itineraries (top 15):**
```{r all-facs-table}
d$all_facs |> head(15) |>
  rename(DETLOC = codes, `Appearances` = n, `% of stays` = pct_of_stays) |> kable()
```

---

## Removal Destinations

The heatmap below crosses **citizenship country** (columns) against **country deported to**
(rows). Cell numbers are absolute counts; cell color shows what share of that row's
removals came from each citizenship group. Top 15 destinations × top 12 citizenships shown.

```{r removal-heatmap, fig.width=12, fig.height=7}
top_dest <- d$removal_heatmap |>
  group_by(departure_country) |>
  summarise(total = sum(n), .groups = "drop") |>
  slice_max(total, n = 15) |>
  arrange(desc(total)) |>
  pull(departure_country)

top_cit <- d$removal_heatmap |>
  group_by(citizenship_country) |>
  summarise(total = sum(n), .groups = "drop") |>
  slice_max(total, n = 12) |>
  arrange(desc(total)) |>
  pull(citizenship_country)

hm <- d$removal_heatmap |>
  filter(departure_country %in% top_dest, citizenship_country %in% top_cit) |>
  tidyr::complete(
    departure_country  = top_dest,
    citizenship_country = top_cit,
    fill = list(n = 0L, row_pct = 0)
  ) |>
  mutate(
    departure_country   = factor(departure_country,  levels = rev(top_dest)),
    citizenship_country = factor(citizenship_country, levels = top_cit)
  )

ggplot(hm, aes(x = citizenship_country, y = departure_country, fill = row_pct)) +
  geom_tile(color = "white", linewidth = 0.5) +
  geom_text(
    aes(label = if_else(n > 0L, format(n, big.mark = ","), "")),
    size = 3, color = "black"
  ) +
  scale_fill_gradient(
    low = "#fff5f0", high = "#a50f15",
    name = "% of row", limits = c(0, 100), na.value = "grey95"
  ) +
  scale_x_discrete(position = "top") +
  labs(
    x = "Citizenship country",
    y = "Deported to",
    title = paste("Citizenship vs. removal destination —", state_full)
  ) +
  theme_minimal(base_size = 11) +
  theme(
    axis.text.x    = element_text(angle = 40, hjust = 0, size = 9),
    axis.text.y    = element_text(size = 9),
    panel.grid     = element_blank(),
    legend.position = "right"
  )
```

**Summary table:**
```{r removal-countries}
d$removal_countries |>
  head(20) |>
  rename(Country = departure_country, N = n, `%` = pct) |>
  kable(caption = paste("Removal destinations for", state_full, "arrests"))
```

---

## Minor Detainees (Under 18)

```{r minors-setup}
m <- d$minors
```

**`r m$n_arrests` arrests** · **`r m$n_persons` distinct individuals** ·
**`r m$n_stays` detention stays.**
Age is estimated as year(arrest) − birth year and may be off by up to one year.

```{r minors-overview}
tibble(Metric = c("Arrests of minors", "Distinct minors", "Detention stays"),
       Value  = c(format(m$n_arrests, big.mark=","),
                  format(m$n_persons, big.mark=","),
                  format(m$n_stays,   big.mark=","))) |> kable()
```

:::{.grid}
:::{.g-col-6}
**Age at arrest**
```{r minors-age}
m$age_dist |> rename(`Age` = age, N = n) |> kable()
```

**Nationality**
```{r minors-nat}
m$nationalities |>
  rename(Country = citizenship_country, N = n, `%` = pct) |> kable()
```
:::
:::{.g-col-6}
**First detention facility**
```{r minors-first}
m$first_fac |>
  rename(DETLOC = code, Facility = facility, N = n, `%` = pct) |> kable()
```

**Last detention facility**
```{r minors-last}
m$last_fac |>
  rename(DETLOC = code, Facility = facility, N = n, `%` = pct) |> kable()
```

**Outcomes**
```{r minors-outcomes}
m$outcomes |>
  rename(`Release reason` = stay_release_reason, N = n, `%` = pct) |> kable()
```
:::
:::

---

## Data notes

- Source: [Deportation Data Project](https://deportationdata.org), FOIA 2026-ICLI-00005.
- Filtered to `apprehension_state == "`r params$state`"` and
  `apprehension_date >= `r params$cutoff``. Duplicate-flagged rows excluded.
- Stays linked via `unique_identifier`; only stays beginning on or after
  `r params$cutoff` are included.
- Transfer flows show consecutive facility pairs from `detention_facility_codes_all`;
  min threshold = `r max(5L, as.integer(d$n_stays * 0.001))` transfers.
- County choropleth coverage: `r round(sum(d$county_counts$n)/d$n_arrests*100,0)`%
  of arrests mapped to a county.
- `stay_release_reason` is NA for stays open as of March 2026 (shown as "Pending / Unknown").
- EROFCB = ERO El Paso Camp East Montana (militarized facility in Montana operated
  by El Paso ERO).
